### 2012-05-28 2011-sgr/ads_defns.tex

\newcommand{\aj}{AJ}                   % Astronomical Journal
\newcommand{\araa}{ARA\&A}             % Annual Review of Astron and Astrophys
\newcommand{\apj}{ApJ}                 % Astrophysical Journal
\newcommand{\apjl}{ApJ}                % Astrophysical Journal, Letters
\newcommand{\apjlett}{ApJ}
\newcommand{\apjs}{ApJS}               % Astrophysical Journal, Supplement
\newcommand{\apjsupp}{ApJS}     
\newcommand{\ao}{Appl.~Opt.}           % Applied Optics
\newcommand{\applopt}{Appl.~Opt.} 
\newcommand{\apss}{Ap\&SS}             % Astrophysics and Space Science
\newcommand{\aap}{A\&A}                % Astronomy and Astrophysics
\newcommand{\astap}{A\&A}  
\newcommand{\aapr}{A\&A~Rev.}          % Astronomy and Astrophysics Reviews
\newcommand{\aaps}{A\&AS}              % Astronomy and Astrophysics, Supplement
\newcommand{\azh}{AZh}                 % Astronomicheskii Zhurnal
\newcommand{\baas}{BAAS}               % Bulletin of the AAS
\newcommand{\jrasc}{JRASC}             % Journal of the RAS of Canada
\newcommand{\memras}{MmRAS}            % Memoirs of the RAS
\newcommand{\mnras}{MNRAS}             % Monthly Notices of the RAS
\newcommand{\pra}{Phys.~Rev.~A}        % Physical Review A: General Physics
\newcommand{\prb}{Phys.~Rev.~B}        % Physical Review B: Solid State
\newcommand{\prc}{Phys.~Rev.~C}        % Physical Review C
\newcommand{\prd}{Phys.~Rev.~D}        % Physical Review D
\newcommand{\pre}{Phys.~Rev.~E}        % Physical Review E
\newcommand{\prl}{Phys.~Rev.~Lett.}    % Physical Review Letters
\newcommand{\pasp}{PASP}               % Publications of the ASP
\newcommand{\pasj}{PASJ}               % Publications of the ASJ
\newcommand{\qjras}{QJRAS}             % Quarterly Journal of the RAS
\newcommand{\skytel}{S\&T}             % Sky and Telescope
\newcommand{\solphys}{Sol.~Phys.}      % Solar Physics
\newcommand{\sovast}{Soviet~Ast.}      % Soviet Astronomy
\newcommand{\ssr}{Space~Sci.~Rev.}     % Space Science Reviews
\newcommand{\zap}{ZAp}                 % Zeitschrift fuer Astrophysik
\newcommand{\nat}{Nature}              % Nature
\newcommand{\iaucirc}{IAU~Circ.}       % IAU Cirulars
\newcommand{\aplett}{Astrophys.~Lett.} % Astrophysics Letters
\newcommand{\apspr}{Astrophys.~Space~Phys.~Res.}
                % Astrophysics Space Physics Research
\newcommand{\bain}{Bull.~Astron.~Inst.~Netherlands} 
                % Bulletin Astronomical Institute of the Netherlands
\newcommand{\fcp}{Fund.~Cosmic~Phys.}  % Fundamental Cosmic Physics
\newcommand{\gca}{Geochim.~Cosmochim.~Acta}   % Geochimica Cosmochimica Acta
\newcommand{\grl}{Geophys.~Res.~Lett.} % Geophysics Research Letters
\newcommand{\jcp}{J.~Chem.~Phys.}      % Journal of Chemical Physics
\newcommand{\jgr}{J.~Geophys.~Res.}    % Journal of Geophysics Research
\newcommand{\jqsrt}{J.~Quant.~Spec.~Radiat.~Transf.}
                % Journal of Quantitiative Spectroscopy and Radiative Transfer
\newcommand{\memsai}{Mem.~Soc.~Astron.~Italiana}
                % Mem. Societa Astronomica Italiana
\newcommand{\nphysa}{Nucl.~Phys.~A}   % Nuclear Physics A
\newcommand{\physrep}{Phys.~Rep.}   % Physics Reports
\newcommand{\physscr}{Phys.~Scr}   % Physica Scripta
\newcommand{\planss}{Planet.~Space~Sci.}   % Planetary Space Science
\newcommand{\procspie}{Proc.~SPIE}   % Proceedings of the SPIE
%
%\let\astap=\aap
%\let\apjlett=\apjl
%\let\apjsupp=\apjs
%\let\applopt=\ao