### 2012-12-03 2012-oss-mrs/Response_to_referee_report.txt
Re-submitted version after GDC and SK comments

Manuscript: LET30039
Title: Hunting the Orphan Stream Host: Identifying Stream Members from Low-Resolution Spectroscopy
Authors: Andrew Casey, Stefan Keller, Gary Da Costa, and Elizabeth Maunder

Dear Mr. Casey,

Your paper above was sent for review to an expert referee who has returned the comments that are appended. The referee has a number of concerns about your paper in its present form and makes many recommendations for improvement in the manuscript. The referee recommends that this paper should be considered for Part 1 of the ApJ, instead of Letters. These suggestions appear to be reasonable to me and I urge you to take the referee's comments fully into account in preparing a revised version of your paper. I will assume that your revised paper will be intended for the main ApJ, so do not worry about the Letters page limit. Letters is able to accept papers into Part 1, so do send the revision back to us for completing the review process.

When you send a revised manuscript, please outline the revisions you have made in response to the referee's comments in a cover letter. Please submit your response in plain text in the field provided when you upload the revised paper. Citing each referee's comment immediately followed by your response to that particular comment would be particularly helpful.

Sincerely,

Fred Rasio
---
Frederic A. Rasio
Letters Editor
The Astrophysical Journal
Northwestern University
ApJLfred@northwestern.edu

====================================================
Referee's report:

This paper may well become a useful contribution to the field of Galactic structure and formation, as it identifies several giant stars that could be members of the Orphan Stream. These stars could be used by future investigators to model the orbit of the stream, identify the progenitor, and possibly the progenitor's extended star formation history. The paper is well written and concise, and I expect it could become a valuable additional resource for other workers in the field.

On the other hand, there are some issues and disagreements with previous work that need to be addressed if we are to have any confidence in the membership probabilities provided in Table 1. In particular, the mean velocity of the stream (as measured using the high probability members) disagrees quite significantly with the measurements of Newberg et al, who made use of a much larger and more widely distributed sample. Of less concern, though still worrisome, is the fact that the current metallicity distribution looks more like that of Newberg et al.'s off-stream, field star sample than it does of the Orphan stream.

Finally, it is not clear to me that this paper needs to be published quickly, and the manuscript already appears to be too long for ApJ Letters. As it stands, the paper is somewhat terse, and with the addition of some explanatory text that I'm recommending (see below), I suspect it will be well over the 4-page limit when ready for publication. I would therefore suggest that the paper be considered for Part 1 of the ApJ instead.

I look forward to seeing a revised version of this manuscript when it becomes available.

Detailed comments follow.

Title

1) ``Host' is not the usual way of referring to a stream's progenitor. Since stars in the stream are unbound from the progenitor and separated by kiloparsecs and billions of years, it is better to say ``Origin of the Orphan Stream' or ``Parent of the Orphan Stream' or ``Orphan Stream Progenitor'.

Response:
Altered title to "Hunting the Parent of the Orphan Stream: Identifying Stream Members from Low-Resolution Spectroscopy", and two instances in the text where the Orphan Stream progenitor was referred to as the host:

 - Section 1
   "as well as make predictions about the nature of the host system."

 - Section 5,
   "help determine both the nature of the host system before it's discovered"

In both cases, 'host system' has been changed to 'progenitor'.

2) In the second sentence, the statement ``left behind in a stream of stars' is somewhat misleading, since tidal streams will extend both forward along the progenitor's orbit as well as behind. This should be amended to something like ``are strewn along the progenitor's orbit in leading and trailing streams of stars.'

Response:
Manuscript edited as requested. Following sentence edited to minimise repeated words. Text now reads as 'Stars which were once gravitationally bound to the satellite are distributed along the progenitor's orbit in leading and trailing streams of stars. The kinematics of the stream..'

3) In the first sentence of the second paragraph and elsewhere in the text, an incorrect reference is made to Grillmair and Dionatos (2006). The correct reference is Grillmair 2006 (ApJ,645, L37).

Response:
Corrected.

Section 2.

4) In the last paragraph, it would better read ``recover stream giants and contaminating foreground dwarfs.'

Response:
Text altered as requested.

5) In this same paragraph, you should specify why no 2mass photometry is available. One assumes this is because even the giants are too faint to be detected in the 2mass survey.

Response:
One assumes correctly. Text altered to read 'Given our colour selection we expect to recover giants and contaminating dwarf contaminants. Although the 2MASS $JHK$ colours can help to separate dwarfs and giants, our target K-giants stars are too faint to be detected in the 2MASS catalogue.'

Section 3

6) ``resolving power' should be ``spectral resolution'.

Response:
Corrected.

7) ``median combined' should be hyphenated.

Response:
Corrected.

8) ``criterion' should be ``criteria'.

Response:
Corrected.

Section 4.2

9) In the first paragraph, it is not clear from Figure 2 why this particular separation line is appropriate or optimal. More explanation
and references should be provided. Also, ``allowing multiple inclusive cuts' is somewhat confusing. Should this be ``imposing simultaneous 
cuts on several observables'?

We have added additional text explaining Figure 2, and why the adopted separation line is appropriate for this work.

``allowing multiple inclusive cuts'' was confusing. The final text for this paragraph now reads as '..we are also employing selections on multiple observables (kinematics, metallicity, proper motions) in order to refine our Orphan Stream giant sample.'

10) In the last paragraph, the selection criteria used by Casey et al. (2012a) should be briefly summarized here.

Response:
The text now reads: 'Given the target selection, our sample is likely to contain many more dwarfs than giants (e.g. see \citet{Casey_et-al_2012} where a similar colour selection was employed).' 
Section 4.3

11) ``fainted' should be ``fainter'.

Response:
Corrected this typo.

12) In Table 1, the ra and dec components of proper motion should be listed separately. Strong cross-stream proper motion components could 
also be used to remove or reduce membership probabilities for some stars.

Response:
The proper motion components in Table 1 are now listed separately. It seems there was an error in the original calculation for the total proper motion of OSS-1. Instead of the proper motion being added in quadrature, the uncertainties were added in quadrature for both the measurement and it's uncertainty (yielding a $mu_{tot} = 6.1 \pm 6.1$). The proper motions are actually (4.1 +/- 4.5, -34.0 +/- 4.5) in RA and Dec respectively. As OSS-1 was already classified as a low-probability member based on inconsistent metallicities between Ca line strengths and isochrone fitting, this makes no change to the probability assignment.

13) Given the small number and significant color and magnitude spread among the highly probable members about the best-fitting isochrone in 
Figure 4, a distance uncertainty of only 1 kpc seems unlikely. How was this uncertainty estimated?

Response:
The distance uncertainty is now determined from propagating the metallicity uncertainty (0.2 dex). A revised distance uncertainty of 2.0 kpc is now reflected throughout the text.

14) The last sentence should include ``21.4 +/- 1 kpc estimated by Grillmair (2006) and Newberg et al. (2010).', since the distance estimates 
in these papers were independent and complementary, one using main sequence fitting and the other using BHB stars.

Response:
Text altered to 'Our derived distance is in reasonably good agreement with the $21.4 \pm 1.0$\,kpc independently deduced by \citet{Grillmair_2006} and \citet{Newberg_et-al_2010}.'

Section 5

15) In the first paragraph, a reference should be given to support the contention that the metallicity spread is consistent with what one might expect for a dwarf galaxy.

Response:
Reference added.

16) The last sentence of the first paragraph should also refer to
Grillmair (2006) main sequence fitting distance measurement.

Response:
Reference rightly added.

17) Given the title of the paper, does the new metallicity measurement help in any way to identify a possible progenitor? At the very least, 
a statement should be included to the extent that the orbit and the metallicity measurements rule out any of the classical dwarfs, and that 
if a bound progenitor still exists, then it must be an as yet undiscovered ultrafaint dwarf.

Response:
Added text: 'Given the stream orbit derived by \citet{Newberg_et-al_2010}, they excluded all possible known halo objects except for the dissolved star cluster, Segue 1. \citet{Simon_et-al_2011} obtained spectroscopy for six members in Segue 1 and found an extremely wide metallicity dispersion: from $<-3.4$ to $-1.63$ dex. On the basis of the extremely low metallicity in the cluster and the wide chemical dispersion, they conclude that Segue 1 is a disrupted dwarf spheroidal galaxy. Although the data presented here indicates the Orphan Stream progenitor is a disrupted dwarf spheroidal galaxy, we cannot reliably associate Segue 1 as the parent without additional observational data.'

18) The maximum likelihood estimates of the stream velocity given by the authors are either 85.3 /pm 4.4 km/s, or 82.1 /pm 1.4 km/s. A simple
weighted mean of the values in Table 1 gives 83.0 km/s. This is very significantly different from the 105 km/s (using the author's V_gsr system) 
measured by Newberg et al. for the same part of the stream using BHB stars. This disagreement needs to be addressed, as it seriously undermines 
any confidence in the purity of even the ``high probability' sample.

Response:
This disagreement was of concern to the authors before the initial manuscript was submitted. We believe this discrepancy is now resolved. We quoted the galactocentric velocity of the stream to be 110 km/s as per Newberg et al. 2011. This was somewhat misleading. Newberg et al. 2011 quotes the mean galactocentric velocity in their conclusions, where this number was originally taken from. However, in Table 2 of Newberg et al. 2011 they find a galactocentric velocity of 101.4 km/s at Lambda_Orphan = 18.4 degrees. This is the closest detection to our fields, at Lambda_Orphan ~= 23 degrees. The referee will recall that this is ~95 km/s on our galactocentric velocity scale. There is a sharp velocity gradient along the stream, decreasing to 38.4 km/s at Lambda_Orphan = 36 degrees. Therefore, at Lambda_Orphan ~= 23 degrees, a galactocentric velocity of 85-90 km/s (or 80-85 km/s on our scale) is quite reasonable. This 'back of the envelope' calculation ignores the intrinsic uncertainty in BHB velocities, or the uncertainties quoted by Newberg et al 2011 of 101.4 km/s +/- 2.9 +/- 9.8 km/s.

Since this implies our initial kinematic selection should actually be between 65 <= V_GSR <= 125 km/s (on our scale), this required a re-analysis of stars in the V_GSR range between 65-75 km/s. No candidates had velocities between 125-135 km/s, thus none were excluded. We identified two more possible candidates that also met our dwarf/giant constraint, and measured their metallicities. These are medium-probability candidates (OSS 15, 20) and have been added to Table 1. Additional explanatory text has been added to the manuscript explaining the velocity results we find in comparison with Newberg et al in Section 4.4
We are glad this error has been spotted and corrected.

19) Moreover, though it is somewhat difficult to say given small number statistics, the metallicity distribution of the high probability stars 
looks more similar to the that of the off-stream, field star sample of Newberg et al. than it does to that of the stream sample.

20) Newberg et al. find velocity dispersions in various parts of the Orphan stream of around 10 km/s. Here the case is made for a much lower 
velocity dispersion, even colder than that of globular cluster streams like Pal 5 and GD-1. It may well be that Newberg et al.'s velocities are
generally less accurate than the present sample. However, one would still have to think about how a stream that is 8-10 times broader than these 
globular cluster streams should be kinematically just as cold or colder. Unfortunately, the disagreement with previous Orphan velocity measurements,
and the fact that removing the weak signal stars makes this disagreement even larger while making the stream colder, undermines my confidence considerably.

21) In thinking about this, I have made use of a velocity versus metallicity plot, with appropriate uncertainties. This makes it much easier to 
see why the authors are getting the results they see, and I would strongly recommend including such a plot in the final manuscript.

Response to (19), (20), (21):
Two additional plots have been added to the manuscript. The first is a plot showing the velocity and metallicity of our highly probable candidates. The second is a histogram showing the metallicity distribution function for our sample, with a comparison to the [Fe/H] of BHB stars in Newberg et al. 2011. The velocity discrepancy with Newberg et al. (2011) has been resolved in (18), and additional discussion text has been added addressing the metallicity differences in Section 4.4.

22) The last sentence of the 3rd paragraph is somewhat mysterious. Is it the random motions that must be perpendicular to the line of sight, or 
the mean motions? The latter follows from the simple fact that the stars are in a stream, and the stream is nearly perpendicular to our line of 
sight. Unless I'm missing something important, this sentence should be deleted.

Response:
Removed the last sentence. It was a bit ambiguous. Sorry about that. AC

----------------------------------------------------------------------