### 2012-12-04 2012-oss-mrs/Response_to_referee_report.txt
Updated resposnse to OSS MRS referee report

Manuscript: LET30039
Title: Hunting the Orphan Stream Host: Identifying Stream Members from Low-Resolution Spectroscopy
Authors: Andrew Casey, Stefan Keller, Gary Da Costa, and Elizabeth Maunder

Dear Mr. Casey,

Your paper above was sent for review to an expert referee who has returned the comments that are appended. The referee has a number of concerns about your paper in its present form and makes many recommendations for improvement in the manuscript. The referee recommends that this paper should be considered for Part 1 of the ApJ, instead of Letters. These suggestions appear to be reasonable to me and I urge you to take the referee's comments fully into account in preparing a revised version of your paper. I will assume that your revised paper will be intended for the main ApJ, so do not worry about the Letters page limit. Letters is able to accept papers into Part 1, so do send the revision back to us for completing the review process.

When you send a revised manuscript, please outline the revisions you have made in response to the referee's comments in a cover letter. Please submit your response in plain text in the field provided when you upload the revised paper. Citing each referee's comment immediately followed by your response to that particular comment would be particularly helpful.

Sincerely,

Fred Rasio
---
Frederic A. Rasio
Letters Editor
The Astrophysical Journal
Northwestern University
ApJLfred@northwestern.edu

====================================================
Referee's report:

This paper may well become a useful contribution to the field of Galactic structure and formation, as it identifies several giant stars that could be members of the Orphan Stream. These stars could be used by future investigators to model the orbit of the stream, identify the progenitor, and possibly the progenitor's extended star formation history. The paper is well written and concise, and I expect it could become a valuable additional resource for other workers in the field.

On the other hand, there are some issues and disagreements with previous work that need to be addressed if we are to have any confidence in the membership probabilities provided in Table 1. In particular, the mean velocity of the stream (as measured using the high probability members) disagrees quite significantly with the measurements of Newberg et al, who made use of a much larger and more widely distributed sample. Of less concern, though still worrisome, is the fact that the current metallicity distribution looks more like that of Newberg et al.'s off-stream, field star sample than it does of the Orphan stream.

Finally, it is not clear to me that this paper needs to be published quickly, and the manuscript already appears to be too long for ApJ Letters. As it stands, the paper is somewhat terse, and with the addition of some explanatory text that I'm recommending (see below), I suspect it will be well over the 4-page limit when ready for publication. I would therefore suggest that the paper be considered for Part 1 of the ApJ instead.

I look forward to seeing a revised version of this manuscript when it becomes available.

Detailed comments follow.

Title

----------------------------------------------------------------------