### 2013-01-30 2012-aqs/draft-aqs.tex
yolo

%!TEX TS-program = pdflatex
\documentclass{emulateapj}

\usepackage{amssymb,graphics} % for resolution symbol
\usepackage{longtable}

\shorttitle{No Evidence for the Aquarius Co-Moving Group Resulting from a Disrupted Classical Globular Cluster}
\shortauthors{Casey et al}

\begin{document}

\title{No Evidence for the Aquarius Co-Moving Group \\Resulting from a Disrupted Classical Globular Cluster\altaffilmark{1}}

% TODO:

% [DONE] Na/O subplots
% [DONE] Na/O comparison with halo
% [DONE] Mg/Al subplots
% - Ni/Na plane plot
% - Add standard stars to observations table
% - Add standard stars chemical abundances
% - Add in-line text referencing standard star chemical abundances
% - Uncertainties analysis
% [DONE] Equivalent widths significant figures.
% - Strontium?
% - Carbon?
% - Distance calculations revision
% - Transverse motions table
% - Transverse motions text
% - Orbital energies plot
% - Minor merger text
% - Conclusions

% Casey, Keller, Alves-Brito, Frebel, Da Costa, Schlaufman, Karakas, Heather, QinQin, Freeman

\author{Andrew R. Casey\altaffilmark{2,3} and distinguished collaborators}
\altaffiltext{1}{This paper includes data gathered with the 6.5 meter Magellan Telescopes located at Las Campanas Observatory, Chile.}
\altaffiltext{2}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia; acasey@mso.anu.edu.au}
\altaffiltext{3}{Massachusetts Institute of Technology, Kavli Institute for Astrophysics and Space Research,
77 Massachusetts Avenue, Cambridge, MA 02139, USA}

\begin{abstract}
We present a detailed analysis of five Aquarius stream candidate stars observed using the MIKE spectrograph with the Magellan Clay telescope.  Enhanced [Na/Fe] and [Ni/Fe] abundance ratios suggest the co-moving group is not the result of a dwarf satellite galaxy, and that its chemical enrichment environment is more representative of either a globular cluster, or the Milky Way thick disc. However, we find no evidence that the stream is a result of a disrupted cluster. Our data indicate the Aquarius stream has a  metallicity dispersion of $\sigma(\mbox{[Fe/H}]) = 0.32$\,dex. No inverse Na-O correlation is present, and we find a positive Mg-Al relationship. Chemically, the group is indistinguishable from the Milky Way thick disk with the exception of one very interesting candidate, which we hypothesise is tidal debris from $\omega$-Centurai. Assuming all other candidates are Aquarius members, we propose the co-moving group has arisen from a minor merger in the Milky Way's thick disk.

\end{abstract}

\keywords{Galaxy: halo, structure --- Individual: Aquarius Stream --- Stars: FGK-giants}

\section{Introduction}
Galaxies are hierarchically formed through chaotic mergers of smaller systems. The Milky Way is no exception. The accumulating stellar debris in our own galactic halo provides ongoing evidence for such merging events. As accreting satellites infall onto the galaxy, tidal forces disrupt the dwarf, hurtling stars in leading and trailing directions. Unsurprisingly, the velocities of stars within these `stellar streams' are sensitive to the galactic potential. As such, their phase-space information can collectively constrain the fraction and distribution of accreted matter in the galaxy, the sub-halo mass function, as well as the shape and extent of dark matter in the Milky Way. Moreover, individual chemical abundances can constrain the  precise nucleosynthesis sites where elements are forged, as well as trace the chemical evolution of the galaxy.

% Consider adding a final sentence about the amount of work that has gone into stellar streaams

Wide-field surveys have proved excellent sources for finding stellar streams. Dozens have been identified through photometric selections and matched-filtering techniques, with some to a galactocentric distance of 100\,kpc \citep{belokurov;et-al_2007,drake;et-al_2013}. Indeed, it is clear that a large fraction of the stellar halo has been built up by accretion. However, as \citet{helmi;white_1999} point out, these observing strategies are only successful for identifying streams that are sufficiently distant from the solar neighbourhood. A nearby stream, within $\sim$10\,kpc, will not appear as an over-density, as the stars would be sparsely positioned across the sky. Such substructures would only be detectable by their kinematics, or perhaps precise chemical abundances through a `chemical tagging' approach (e.g. see \citet{freeman;bland-hawthorn_2002}).

% Talk about RAVE
It is therefore necessary to survey solar neighbourhood stars with spectroscopy to reveal any nearby substructures. The RAVE (Radial Velocity Experiment) team began such a survey in 2003 and have observed spectra from over 500,000 stars across 17,000\,deg$^{2}$. As the name hints, the primary goal of RAVE was to obtain radial velocities for the solar neighbourhood and beyond. In an attempt to remain kinematically unbiased, RAVE candidates were selected based only on their apparent magnitude. With the exception of bulge stars, where some photometric cuts were employed, all stars between 9 $<$ I $<$ 13 were observed. Almost all of these candidates have published radial velocity measurements \citep{steinmetz;et-al_2006}, and for a subset of stars with sufficient Signal-to-Noise ($S/N$), estimates of stellar parameters have been derived using a $\chi^2$ minimization technique \citep{zwitter;et-al_2008, siebert;et-al_2011}.

% Identification of the aquarius stream
Using these data, \citet{williams;et-al_2011} identified a co-moving group of nearby ($0.5\,\mbox{kpc} \lesssim d \lesssim 10\,\mbox{kpc}$) stars in the vicinity of the Aquarius constellation $(l, b) = (60\,^\circ, -55\,^\circ)$. Thus, the co-moving group was named the Aquarius stream. The artefact is most apparent when examining velocities against galactic latitude for stars within $-70\,^\circ < b < -50\,^\circ$. \citet{williams;et-al_2011} employed a selection criteria of $-250\,\mbox{km s}^{-1} < V_{\rm hel} < -150\,\mbox{km s}^{-1}$, $30\,^\circ < l < 75\,^\circ$ and $J > 10.3$ to maximize the contrast between the stream and background, identifying 15 candidates in the process. The average heliocentric velocity of these members was found to be $V_{\rm hel} = -199\,\mbox{km s}^{-1}$, with a dispersion of 27\,km s$^{-1}$. The radial velocity uncertainties provided by the RAVE catalog are described to be $\sim$2\,km s$^{-1}$, so the wide kinematic distribution appears to be real.

% QUANTIFYING THE STREAM DETECTION
Upon examining the galactic longitude and velocities of Aquarius stars, the stream appears quite distinct from the halo. In order to quantify the statistical likelihood that the Aquarius stream is truly separate from a smooth halo distribution, \citet{williams;et-al_2011} compared the number of observed RAVE stars within $-70\,^\circ < b < -50\,^\circ$ against the Galaxia \citep{sharma;et-al_2011} and Besan\c{c}on \citep{robin;et-al_2003} galaxy simulations. The predicted particles in each model between $-70\,^\circ < b < -50\,^\circ$ were binned in $\Delta{l} \times \Delta{V_{\rm hel}}$ blocks using varying grid cell sizes between $25\,^\circ$ to $70\,^\circ$ and $20\,\mbox{km s}^{-1}$ to $100\,\mbox{km s}^{-1}$. The RAVE sample was binned in the same way, and the statistical significance of the observed over-density is discretized. \citet{williams;et-al_2011} found the stream to be statistically significant ($>$4-$\sigma$) in 96\% of the cell proportions employed for the Besan\c{c}on model. By sampling the Galaxia model repeatedly with varying extinction rates, a similar result was found: between $75\pm15$\,\% to $80\pm18$\,\% of grid cells showed the Aquarius stream to significant at the 4-$\sigma$ level or higher. Given the high-latitude and proximity of the stream, it is not surprising that dust had a negligible effect on the statistical significance. \citet{williams;et-al_2011} concluded that the over-density in $l, V_{\rm hel}$ is a true substructure, and neither the choice of galaxy model, cell dimension, nor the extinction model made any real difference to the detection significance.

% STREAM IS LOCALIZED
% the stream appears to be quite localized in space.
% nearby streams could be strewn out over several degrees. the Orphan stream wraps over 60 degrees in the sky, and the Sgr stream wraps several steradians

% DISTANCE AND DYNAMICS
% heliocentric + galactic velocities suggest the aquarius stream stars are in the halo, and not associated with the disk.
% distances

The Sagittarius stream is arguably the most prominent ongoing accretion event in the galaxy. Given the extent of stellar debris the dwarf has littered throughout the Milky Way, it is reasonable to suspect the Aquarius stream might originate from the Sagittarius dwarf. Indeed, the Aquarius  stars do not lie far from the orbital plane of Sagittarius, and the metallicities reported by \citet{williams;et-al_2011} are not dissimilar from the Sagittarius stream. However, \citet{williams;et-al_2011} found the Aquarius transverse motions (specifically $V_{Z}$ and $V_\phi$) to be quite distinct from Sagittarius. Based on the phase space information available, \citet{williams;et-al_2011} concluded that the newly discovered stream could not be positively associated with the Monoceros stream, Hercules-Aquila cloud, or either the Canis Major or Virgo over-densities.

The Aquarius stream has a metallicity of $\mbox{[Fe/H]} \sim -1.0$, slightly more metal-rich than halo stars at the same distance. \citet{williams;et-al_2011} found the stream to have an average of $\mbox{[M/H]} = -1.0 \pm 0.4$, compared to $\mbox{[M/H]} = -1.1 \pm 0.6$ for background stars after the same selection cuts had been employed. Of the 12 Aquarius stars in the \citet{williams;et-al_2011} discovery sample with stellar parameters, the metallicity range is wide: between --2.02 and --0.33 dex. Spectroscopic observations of high resolution with high $S/N$ are necessary to accurately characterize the stream's metallicity distribution.

To this end, \citet{wylie-de-boer;et-al_2012} obtained high-resolution ($\mathcal{R} = 25,000$) spectra with modest $S/N$ ($\sim$30 pixel$^{-1}$) for six Aquarius stream stars. Their data indicate a very narrow metallicity distribution: $\mbox{[Fe/H]} = -1.09 \pm 0.10$\,dex extending between $\mbox{[Fe/H]} = -1.25$ to $-0.98$\,dex. Observations of such minimal chemical scatter from stars with a common origin is typically noted in either globular, or open clusters. The largest [Fe/H] discrepancy between the \citet{williams;et-al_2011} and \citet{wylie-de-boer;et-al_2012} study was $\Delta\mbox{[Fe/H]} = -0.66$\,dex for the most metal-rich star in the \citet{williams;et-al_2011} sample. The most metal-poor star in the \citet{williams;et-al_2011} study was not observed by \citet{wylie-de-boer;et-al_2012}.

In addition to ascertaining stellar parameters, \citet{wylie-de-boer;et-al_2012} measured elemental abundances for the Aquarius stream stars -- the only study to do so until now. The authors primarily focussed on light elements: Na, O, Ni, Mg, and Al. Chemical patterns in these light elements arise as natural consequences of nucleosynthetic processes in massive stars. These patterns have been repeatedly observed in globular clusters. Specifically, an anti-correlation between sodium and oxygen content appears ubiquitous to globular clusters. Originally this relationship was thought to arise from internal mixing along the giant branch. However, dwarf globular cluster stars were found to exhibit strong Na-enhancement and O-depletion, contrary to what would be expected if the Na-O anti-correlation was purely a result of internal mixing. It is well known now that photospheric abundance variations resulting from internal mixing are limited to carbon, nitrogen and lithium. Additionally, light element variations can be explained by classic nucleosynthetic pathways, albeit extreme temperatures are necessary for such relationships to arise \--- much higher than would reasonably be expected in globular cluster stars. Several descriptions for an astrophysical site that would permit such nucleosynthetic processes have been hypothesized, including massive intermediate-mass Asymptotic Giant Branch (AGB) binary stars \citep{de_mink;et-al_2009} and fast-rotating massive stars \citep{decressin;et-al_2007}. While the exact formation site is still under debate, the chemical relationship is well-established empirically. The Na-O pattern provides the best chemical evidence that their chemical enrichment history has been dominated by Sn \textsc{ii} events and that those stars formed in a globular cluster-like environment.

From high-resolution spectroscopy of five candidate Aquarius stars, four of which had sodium and oxygen abundance measurements, \citet{wylie-de-boer;et-al_2012} found two stars with super-solar sodium; slightly higher [Na/Fe] ratios than halo stars of the same metallicity. However, no strong oxygen depletion was evident in the data. Thus, they observed no evidence of a classical Na-O inverse correlation in the Aquarius stream.

There are other chemical abundance patterns that allow us to infer the chemical enrichment environment. \citet{nissen;schuster_1997} first noted that nearby solar neighbourhood stars were enriched in [Ni/Fe] and [Na/Fe]. \citet{fulbright_2000} came to a similar conclusion: only somewhat distant ($R_\odot > 20$\,kpc) stars are likely to be sodium deficient. If the stellar halo has been built up from the accretion of smaller dwarf satellites, \citet{nissen;schuster_1997} proposed that this chemical trend may be indicative of the halo merger history since stars at larger $R_\odot$ are more likely to have been accreted rather than formed in-situ. \citet{wylie-de-boer;et-al_2012} found all the Aquarius stars in their sample to be enriched in both sodium and nickel; contrary to what would be expected if the stars originated from a disrupted dwarf spheroidal (dSph) galaxy. This provided an additional indication that the Aquarius stream stars originated in a globular cluster-like environment.

Combined with the low level of chemical scatter present in their sample, these chemical indicators led \citet{wylie-de-boer;et-al_2012} to conclude that the Aquarius stream is the result of a tidally disrupted globular cluster. \citet{williams;et-al_2011} had previously precluded this possibility with any known globular clusters after modelling an Aquarius-like progenitor infalling onto the Milky Way and comparing the predicted $l, b, V_{\rm hel}$ with known, nearby globular clusters of similar metallicities. No known globular cluster could match their simulation. If both scenarios are accurate then the undiscovered disrupting cluster might still remain, and would be relatively close by.

We seek to investigate the nature of the Aquarius stream, specifically the recent globular cluster origin claim made by \citet{wylie-de-boer;et-al_2012}. We present a detailed analysis for five Aquarius stream candidates from high-resolution, high $S/N$ observations taken using the Magellan Inamori Kyocera Echelle spectrograph \citep{bernstein;et-al_2003} with the Magellan Clay telescope. Details of the observations and data reduction are outlined in the following section. The data analysis is chronicled in Section \ref{sec:analysis} and a detailed discussion of these results resides in Section \ref{sec:discussion}. In Section \ref{sec:conclusions} we present our conclusions and critical interpretations.

\begin{deluxetable*}{lcccccccccc}
\tablecolumns{1}
\tabletypesize{\scriptsize}
\tablecaption{Observations\label{tab:observations}}
\tablehead{
	\colhead{Designation} &
	\colhead{$\alpha$} &
	\colhead{$\delta$} &
	\colhead{Observed} &
	\colhead{Airmass} &
	\colhead{Seeing} &
	\colhead{$t_{exp}$} &
	\colhead{S/N\tablenotemark{a}} &
	\colhead{$V_{\rm rad}$} &
	\colhead{$V_{\rm hel}$} &
	\colhead{$V_{\rm err}$} \\
 & (J2000) & (J2000) & Date & & (") & (secs) & (px$^{-1}$) & (km s$^{-1}$) & (km s$^{-1}$) & (km s$^{-1}$)
}
\startdata

\\ \multicolumn{11}{c}{Standard Stars} \\ \hline \\
HD 41667		& 06:05:03.7 & $-$32:59:36.8	& 2011-03-13	& 1.005	& \dots & \dots & \dots & 314.4	& \dots & 0.8 \\
HD 142948		& 16:00:01.6 & $-$53:51:04.1	& 2011-03-14	& 1.107	& \dots & \dots & \dots & 6.8		& \dots & 0.4 \\

\\ \multicolumn{11}{c}{Program Stars} \\ \hline \\
C2225316-14437	& 22:25:31.7 & $-$14:54:39.6	& 2011-07-30	& 1.033 & \dots & \dots & 135	 & --168.6	& --156.4 & 0.1 \\
C2306265-085103	& 23:06:26.6 & $-$08:51:04.8	& 2011-07-30	& 1.096 & \dots & \dots & 100	 & --240.4	& --221.1 & 0.1 \\
J221821-183424	& 22:18:21.2 & $-$18:34:28.3	& 2011-07-30	& 1.026	& \dots & \dots & 115 	 & --170.2	& --159.5 & 0.1 \\
J223504-152834	& 22:35:04.5 & $-$15:28:34.9	& 2011-07-30	& 1.047	& \dots & \dots & 135	 & --179.0	& --169.7 & 0.1 \\
J223811-104126	& 22:38:11.6 & $-$10:41:29.4	& 2011-07-30	& 1.218	& \dots & \dots & 115	 & --250.1	& --235.7 & 0.1

% V corrections (already applied the first number of each row)
% J221821     	:   0.30, 0.25 +/- 1.25 (MEK has -154.1, wdb ) -- diff: 5.4  (positive == higher)
% C2225316		:	0.43, 0.42 +/- 1.13 (MEK has -155.7) -- diff: -0.4 (hers is lower)
% C2306265		: 	1.14, 1.19 +/- 1.16 (MEK has -221.8) -- diff: -0.7 (hers is lower)
% J223504		:	-1.87, -1.89 +/- 1.45 (MEK has -166.9) -- diff: 2.8 (positive == higher)
% J223811		:	1.69, 1.61	+/- 1.47 (MEK has --230.1) -- diff 5.6 (positive == higher)

\enddata
\tablenotetext{a}{S/N measured at 6000 \AA{} for each target.}
\end{deluxetable*}

\section{Observations \& Data Reduction}

The most complete sample of Aquarius stream stars is presented in the discovery paper of \citet{williams;et-al_2011}. Our sample includes five stars: four common to the \citet{wylie-de-boer;et-al_2012} study, and an additional candidate from the original \citet{williams;et-al_2011} sample. The supplementary star, J2306265-085103, was observed by the RAVE survey but had a $S/N$ ratio too low for stellar parameters to be accurately determined. All program stars were observed in July 2011 in good seeing at low airmass (Table \ref{tab:observations}), and X additional standard stars were observed in March 2011. All observations were taken using a 1.0'' slit, providing a spectral resolution of $\mathcal{R} = 28,000$ in the blue arm and $\mathcal{R} = 22,000$ in the red arm. The total exposure times for our program stars varied to ensure a $S/N$ ratio in excess of 100\,pixel$^{-1}$ at 600\,nm.

% what spectral resolution in the blue, what resolution in the red

Calibration frames were taken at the start of each night, including twenty flat-field frames (10 quartz, 10 milky) and 10 Th-Ar arc lamp exposures for wavelength calibration. The data were reduced using the CarPy pipeline\footnote{http://code.obs.carnegiescience.edu/mike}. One standard stars, HD 41667, was also reduced using standard extraction and calibration methods in \textsc{iraf}. The resultant spectra from both approaches were compared for residual fringing, $S/N$, and wavelength calibration. No noteworthy differences were present, and the CarPy pipeline was utilized for the remainder of the data reductions. Each reduced echelle order was carefully normalized using a third order spline with defined knot spacing. Normalized orders were stitched together to provide a single one-dimensional spectrum from 3800-9400\,\AA{}.

The white dwarf HR 6141 was observed on X as a telluric standard. The $S/N$ ratio for HR 6141 exceeds that in any of our standard or program stars. Although the atmospheric conditions at Las Campanas Observatory are certain to change throughout the night and between observing runs, we are primarily using this spectra to identify lines that are affected by telluric absorption. Critical transitions necessary for abundance analyses that are affected by the Earth's atmosphere are corrected for telluric absorption.

\section{Analysis}
\label{sec:analysis}

\subsection{Radial Velocities}
\label{sec:radial-velocities}

The radial velocity for each star was determined in a two-step method. An initial estimate of the radial velocity was ascertained by cross-correlation with a synthetic spectrum of a giant star with $T_{\rm eff} = 4500$\,K, $\log{g} = 1.5$, and [M/H] $= -1.0$ across the wavelength range $8450 - 8700$ \AA{}. The observed spectrum was shifted to the pseudo-rest frame using this initial velocity estimate. Velocities found from cross-correlation are typically within 1\,km\,s$^{-1}$ of our final published value.

\begin{figure}[h!]
	\includegraphics[width=\columnwidth]{./figures/line-velocity.pdf}
	\caption{Residual velocities for each absorption line with a well-fitted Gaussian profile after correcting for doppler shift using the initial cross-correlation measurement. The mean residual velocity $-0.75$\,km\,s$^{-1}$ is within the 1-$\sigma$ cross-correlation uncertainty of $\pm0.9$\,km\,s$^{-1}$.}
	\label{fig:line-velocities}
\end{figure}

With our spectra at near-rest, we measured the equivalent widths of $\sim$200 lines by automatically fitting Gaussian functions to the absorption profiles (see Section \ref{sec:line-measurements}). Fitted Gaussian profiles provide us with a measured equivalent width, the full-width half-maximum (FWHM) and the central wavelength of the profile. Thus, the residual radial velocity can be found from the ratio of the central and rest wavelengths. The final radial velocity for the star is provided by the initial velocity and the mean of $\sim$200 residual line velocity measurements. Figure \ref{fig:line-velocities} shows the line velocities for HD 41667 after being placed at pseudo-rest using our cross-correlation velocity of 314.4\,km s$^{-1}$. As expected, the mean residual offset is small ($-0.75$\,km s$^{-1}$), and the standard deviation is 0.79\,km s$^{-1}$ from 164 line measurements. This provides us with a final measured radial velocity of $313.7 \pm 0.1$\,km s$^{-1}$. This process was performed for all observations. The radial velocities published in Table \ref{tab:observations} are the final values from this two-step method. Heliocentric velocities agree quite well with those compiled by \citet{williams;et-al_2011} from the RAVE survey: the mean offset is $\langle\Delta{V_{\rm hel}}\rangle = 2.5 \pm 2.7$\,km\,s$^{-1}$, demonstrating the accuracy of the velocities published in the RAVE catalog.

\subsection{Line Measurements}
\label{sec:line-measurements}

For the measurement of atomic absorption lines, we employed the line list of \citet{yong;et-al_2005} with additional transitions of Cr, Sc, Zn, and Sr from \citet{frebel;et-al_2010}. The line list was augmented with hyperfine-structure data for Sc and Mn from the Kurucz compilation \citet{kurucz;1998}. Molecular line data for CH was taken from \citet{plez;et-al_2008,plez;et-al_2009}. For lines with hyperfine structure, blended transitions or molecular features, we determined the abundance using a spectral synthesis approach. For all other transitions, equivalent widths were measured to determine the line abundance.

The equivalent widths for all absorption lines were measured automatically using software written for this study. Given the rest wavelength and an initial guess of the FWHM, a Gaussian profile is iteratively fitted to the absorption line. At the same time, the local continuum is found within 20\,\AA{} either side of the rest wavelength using a second-order polynomial. Measuring the local continuum ensures any errors in our initial normalisation or order-stitching do not propagate through to our equivalent width measurements. When determining the local continuum, any neighbouring group of pixels that deviate significantly ($>$2-$\sigma$) from the local continua either belong to a persistent cosmic ray hit, or they are trough points of a neighbouring absorption profile. We attempt to fit a Gaussian profile to each group of deviating pixels, using the current best estimate for the local continuum. If a Gaussian profile is successfully fitted to the deviating group and its surrounding pixels, we assume the outlier points belong to an absorption line and all pixels belonging to that profile are excluded from the continuum determination.

When fitting Gaussian profiles, the $\chi^2$ difference between the observed spectra and the fitted function is minimized. In order to account for crowded absorption regions, where the $\chi^2$ value might be influenced by a nearby or blended profile, our $\chi^2$ function is weighted based on the distance to the rest wavelength by the function

\begin{equation}
W_{\lambda_{i}} = 1 + \exp{\left(\frac{-(\lambda_{i} - \lambda_{r})^2}{4\sigma^2}\right)}
\label{eq:chi-weight}
\end{equation}

\noindent where the rest wavelength ($\lambda_{r}$) and standard deviation\footnote{Recall $\rm{FWHM} = 2\sigma\sqrt{2\log_e{2}}$} ($\sigma$) are free parameters in the iterative fitting process. As a consequence of Eq. \ref{eq:chi-weight}, pixels near the rest wavelength are weighted higher than those on the wings, forcing the fitting scheme to disregard blended or crowded surroundings. Although this approach relies on a reasonably accurate initial radial velocity correction of $<$4\,km s$^{-1}$ \--- which is easily achieved \--- it greatly improves the accuracy and robustness of the equivalent width measurements.

The results of our iterative Gaussian fitting approach are \textit{completely insensitive} to the initial FWHM guess. Increasing the initial FWHM estimate from 0.1\,\AA{} to 1 or 2\,\AA{} \--- which are unphysically large values for high-resolution spectra \--- does not alter either the FWHM or the equivalent width for any of our measurements. Only a small increase in computational cost is observed. Although we are extremely confident in our equivalent width measurements, every absorption profile was examined by eye for quality, and spurious measurements were excluded.

We list the atomic data and measured equivalent widths for transitions used during this analysis in Table \ref{tab:equivalent-widths}. Saturated lines were excluded by removing measurements with reduced equivalent widths, $\log_{10}{(EW/\lambda)} > -4.5$\, dex. A minimum detectable equivalent width was calculated as a function of wavelength based on the $S/N$, and only lines that exceeded a 3-$\sigma$ detection significance were included. We have verified the approach by comparing our measurements of 156 lines in HD 140283 with the study of \citet{norris;et-al_1996}. We only included measurements in the \citet{norris;et-al_1996} study that were not marked with questionable line quality parameters. Excellent agreement is found between the two studies, which is illustrated in Figure \ref{fig:ew-compare}. The mean difference is a negligible $-0.64 \pm 2.78$\,m\AA{}, and no systematic trend is present. We note that the scatter may be directly attributed to the observed $S/N$ ratio, as other studies \citep{frebel;et-al_2013} using the same equivalent width measurement software described here find better agreement with hand measurements for spectra with higher $S/N$ ratios: $0.20 \pm 0.16$\,m{\AA} with \citet{aoki;et-al_2007} and $0.25 \pm 0.28$\,m{\AA} with \citet{cayrel;et-al_2004} for a spectrum of HD 122563 with $S/N > 300$\,pixel$^{-1}$.

% equivalent widths plot with Norris for HD 122563
\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./figures/smh-norris.pdf}
	\caption{A comparison showing equivalent widths measured for HD 140283 using our automatic routine (see \S\ref{sec:line-measurements}), and those found by careful hand measurements in \citet{norris;et-al_1996}. No systematic trend is present, and the mean difference between these studies is $\langle\Delta{}W_\lambda\rangle = -0.64 \pm 2.78$ m\AA{}.}
	\label{fig:ew-compare}
\end{figure}

\begin{deluxetable*}{cccccccccc}
\tabletypesize{\scriptsize}
\tablecaption{List of Line Transitions and Equivalent Width Measurements\label{tab:equivalent-widths}}
\tablehead{
	& & & & \multicolumn{5}{c}{Equivalent Width} \\
	\cline{5-9} \\
	\colhead{Wavelength} &
	\colhead{Transition} &
	\colhead{LEP} &
	\colhead{$\log{gf}$} &
	\colhead{C2225316-14437} &
	\colhead{C2306265-085103} &
	\colhead{J221821-183424} &
	\colhead{J223504-152834} &
	\colhead{J223811-104126} \\
	({\AA})	& & (eV) & & (m{\AA}) & (m{\AA}) & (m{\AA}) & (m{\AA}) & (m{\AA}) 
}
\startdata 
 6300.300 &       8.0 &      0.00 &    --9.717 &      45.4 &      66.9 &      38.5 &      32.8 &      17.9 \\
 6363.780 &       8.0 &      0.02 &   --10.185 &      19.5 &      28.3 &      13.0 &      18.8 &   \nodata \\
 5688.190 &      11.0 &      2.11 &    --0.420 &   \nodata &   \nodata &      49.0 &     131.5 &      38.4 \\
 6154.230 &      11.0 &      2.10 &    --1.530 &      24.1 &      38.9 &   \nodata &      48.5 &   \nodata \\
 6160.750 &      11.0 &      2.10 &    --1.230 &      37.7 &      58.5 &   \nodata &      65.7 &   \nodata \\
 6318.720 &      12.0 &      5.11 &    --1.970 &   \nodata &      47.9 &      14.7 &      62.8 &       8.9 \\
 6319.240 &      12.0 &      5.11 &    --2.220 &      30.8 &   \nodata &       5.5 &   \nodata &   \nodata \\
 6965.410 &      12.0 &      5.75 &    --1.510 &   \nodata &   \nodata &   \nodata &      59.2 &   \nodata \\
 7387.690 &      12.0 &      5.75 &    --0.870 &      51.7 &      56.3 &      24.7 &      76.0 &      24.0 \\
 6696.020 &      13.0 &      3.14 &    --1.340 &      72.8 &      63.6 &      13.0 &   \nodata &   \nodata \\
 6698.670 &      13.0 &      3.14 &    --1.640 &      39.7 &      33.4 &   \nodata &      39.5 &   \nodata \\
 7835.310 &      13.0 &      4.02 &    --0.470 &      47.8 &      38.4 &   \nodata &      55.9 &       5.4 \\
 7836.130 &      13.0 &      4.02 &    --0.310 &      62.8 &      44.7 &   \nodata &      69.1 &      13.2 \\
 5690.430 &      14.0 &      4.93 &    --1.751 &      56.2 &      43.8 &      25.7 &      53.4 &      16.8 \\
 5793.070 &      14.0 &      4.93 &    --1.843 &      46.1 &      39.8 &      19.3 &      48.4 &   \nodata \\
 6125.020 &      14.0 &      5.61 &    --1.506 &      40.9 &      23.4 &      10.9 &      34.7 &   \nodata \\
 6145.010 &      14.0 &      5.62 &    --1.362 &      33.9 &      21.7 &      16.1 &      32.7 &      10.0 \\
 6155.130 &      14.0 &      5.62 &    --0.786 &      67.4 &      53.1 &      35.3 &      72.0 &      25.0 \\
 6166.440 &      20.0 &      2.52 &    --1.142 &      94.2 &     105.5 &      42.3 &      99.5 &      35.9 \\
 6169.040 &      20.0 &      2.52 &    --0.797 &     112.1 &     120.9 &      66.3 &     120.5 &      55.5 \\
 6169.560 &      20.0 &      2.53 &    --0.478 &     127.9 &     141.2 &      83.7 &     136.4 &      71.3 \\
 6455.600 &      20.0 &      2.52 &    --1.290 &      80.8 &      91.4 &      31.4 &      89.2 &      26.0 \\
 5526.820 &      21.1 &      1.77 &      0.130 &      98.8 &     113.3 &     100.8 &      98.4 &      51.1 \\
 5657.880 &      21.1 &      1.51 &    --0.500 &      96.5 &     107.5 &      92.7 &      95.4 &      41.2 \\
 5667.150 &      21.1 &      1.50 &    --1.240 &      59.4 &      75.0 &      45.0 &      58.5 &      14.5 \\
 5669.040 &      21.1 &      1.50 &    --1.120 &   \nodata &      92.5 &      57.3 &      75.2 &      18.2 \\
 6245.610 &      21.1 &      1.51 &    --0.980 &      67.1 &      83.6 &      53.3 &      67.4 &      14.8 \\
 6604.600 &      21.1 &      1.36 &    --1.480 &      65.6 &      82.1 &      49.4 &      65.9 &      17.8 \\
 6064.630 &      22.0 &      1.05 &    --1.888 &      59.7 &      80.9 &   \nodata &      58.6 &   \nodata \\
 6091.170 &      22.0 &      2.27 &    --0.367 &      45.2 &      59.1 &   \nodata &      56.3 &   \nodata \\
 6312.240 &      22.0 &      1.46 &    --1.496 &      46.2 &      63.4 &   \nodata &      48.8 &   \nodata \\
 6336.100 &      22.0 &      1.44 &    --1.687 &      35.9 &      54.1 &   \nodata &      42.3 &   \nodata \\
 5154.080 &      22.1 &      1.57 &    --1.920 &     112.2 &   \nodata &     114.0 &   \nodata &      53.9 \\
 5185.910 &      22.1 &      1.89 &    --1.350 &   \nodata &     101.4 &     105.5 &   \nodata &      55.3 \\
 5336.790 &      22.1 &      1.58 &    --1.690 &     116.3 &     124.6 &     119.5 &     104.0 &      61.4 \\
 5381.030 &      22.1 &      1.59 &    --2.080 &   \nodata &     131.4 &     101.9 &     109.9 &      48.0 \\
 5727.060 &      23.0 &      1.08 &    --0.010 &     122.4 &   \nodata &      26.0 &     116.8 &       9.1 \\
 6090.220 &      23.0 &      1.08 &    --0.060 &      93.6 &     115.6 &      23.6 &      92.8 &   \nodata \\
 6216.360 &      23.0 &      0.28 &    --1.290 &     108.8 &   \nodata &      16.9 &     109.3 &   \nodata \\
 6251.820 &      23.0 &      0.29 &    --1.340 &      94.7 &     124.4 &   \nodata &      92.4 &   \nodata \\
 6274.640 &      23.0 &      0.27 &    --1.670 &      65.5 &      93.8 &   \nodata &      61.5 &   \nodata \\
 6504.160 &      23.0 &      1.18 &    --1.230 &      22.9 &      33.7 &   \nodata &      31.9 &   \nodata \\
 4545.950 &      24.0 &      0.94 &    --1.370 &   \nodata &   \nodata &      80.2 &   \nodata &   \nodata \\
 4580.056 &      24.0 &      0.94 &    --1.650 &   \nodata &   \nodata &   \nodata &   \nodata &      44.2 \\
 4616.137 &      24.0 &      0.98 &    --1.190 &     128.7 &     139.3 &      89.4 &     126.2 &      59.9 \\
 4626.188 &      24.0 &      0.97 &    --1.320 &     122.2 &     136.4 &      82.2 &   \nodata &      56.9 \\
 4646.150 &      24.0 &      1.03 &    --0.740 &   \nodata &   \nodata &     110.6 &     146.1 &      75.3 \\
 4651.280 &      24.0 &      0.98 &    --1.460 &     127.8 &     140.6 &      75.4 &     116.3 &      48.4 \\
 4652.158 &      24.0 &      1.00 &    --1.030 &     138.5 &   \nodata &      96.9 &     138.0 &      66.7 \\
 5206.040 &      24.0 &      0.94 &      0.020 &   \nodata &   \nodata &   \nodata &   \nodata &     132.8 \\
 5247.564 &      24.0 &      0.96 &    --1.640 &     141.4 &   \nodata &      76.1 &   \nodata &      46.7 \\
 5296.690 &      24.0 &      0.98 &    --1.360 &     149.9 &   \nodata &      85.0 &     141.0 &      59.0 \\
 5300.740 &      24.0 &      0.98 &    --2.000 &     120.5 &   \nodata &   \nodata &     112.2 &   \nodata \\
 5345.800 &      24.0 &      1.00 &    --0.950 &   \nodata &   \nodata &     115.4 &   \nodata &      81.1 \\
 5348.310 &      24.0 &      1.00 &    --1.210 &     150.8 &   \nodata &      94.8 &     142.4 &      63.0 \\
 5409.770 &      24.0 &      1.03 &    --0.670 &   \nodata &   \nodata &     132.4 &   \nodata &      92.1 \\
 4558.594 &      24.1 &      4.07 &    --0.656 &   \nodata &   \nodata &   \nodata &     101.8 &      46.5 \\
 4588.142 &      24.1 &      4.07 &    --0.826 &      46.0 &      53.0 &      60.5 &      55.0 &      36.0 \\
 4591.992 &      24.1 &      4.07 &    --1.419 &   \nodata &   \nodata &      29.1 &      28.5 &      19.6 \\
 6013.530 &      25.0 &      3.07 &    --0.251 &      89.2 &     114.6 &      22.2 &     109.4 &      14.8 \\
 6016.670 &      25.0 &      3.08 &    --0.100 &      94.0 &     116.0 &      28.1 &     113.0 &      17.1 \\
 6021.800 &      25.0 &      3.08 &      0.034 &      97.5 &     116.0 &      39.1 &     110.7 &      19.9 \\
 5044.210 &      26.0 &      2.85 &    --2.034 &   \nodata &   \nodata &   \nodata &   \nodata &      33.7 \\
 5054.640 &      26.0 &      3.64 &    --1.938 &      38.0 &   \nodata &   \nodata &   \nodata &   \nodata \\
 5242.490 &      26.0 &      3.63 &    --0.980 &      94.3 &     101.8 &      72.9 &   \nodata &      49.2 \\
 5321.110 &      26.0 &      4.43 &    --1.106 &      35.0 &      35.3 &      11.1 &   \nodata &   \nodata \\
 5322.040 &      26.0 &      2.28 &    --2.840 &      93.1 &     103.0 &      59.5 &   \nodata &      30.9 \\
 5326.140 &      26.0 &      3.57 &    --2.130 &      40.4 &      51.7 &      13.7 &      48.4 &   \nodata \\
 5365.400 &      26.0 &      3.57 &    --1.040 &   \nodata &     100.3 &   \nodata &      93.8 &   \nodata \\
 5367.480 &      26.0 &      4.41 &      0.430 &     108.4 &     118.0 &      92.4 &     119.1 &      73.7 \\
 5379.570 &      26.0 &      3.69 &    --1.530 &      71.4 &      81.4 &      34.9 &      75.6 &      19.6 \\
 5491.840 &      26.0 &      4.18 &    --2.250 &   \nodata &   \nodata &   \nodata &      17.3 &   \nodata \\
 5618.630 &      26.0 &      4.21 &    --1.292 &      43.5 &      54.3 &      20.2 &      57.4 &      11.3 \\
 5701.550 &      26.0 &      2.56 &    --2.216 &     118.5 &     134.8 &      91.8 &     119.5 &      49.5 \\
 5705.470 &      26.0 &      4.30 &    --1.420 &      29.2 &      41.5 &      10.9 &      44.5 &       7.7 \\
 5741.850 &      26.0 &      4.25 &    --1.689 &      36.2 &      37.0 &      13.0 &      48.2 &   \nodata \\
 5775.080 &      26.0 &      4.22 &    --1.310 &      55.0 &      62.1 &      25.3 &      67.4 &      14.7 \\
 5778.450 &      26.0 &      2.59 &    --3.480 &      42.8 &      55.8 &      13.4 &      50.1 &   \nodata \\
 5811.920 &      26.0 &      4.14 &    --2.430 &   \nodata &   \nodata &   \nodata &      10.9 &   \nodata \\
 5837.700 &      26.0 &      4.29 &    --2.340 &   \nodata &   \nodata &   \nodata &      12.9 &   \nodata \\
 5853.160 &      26.0 &      1.49 &    --5.280 &      30.4 &      43.4 &       6.3 &      34.7 &   \nodata \\
 5855.090 &      26.0 &      4.60 &    --1.547 &      13.3 &      20.3 &   \nodata &      22.8 &   \nodata \\
 5856.100 &      26.0 &      4.29 &    --1.640 &      27.9 &      40.2 &   \nodata &      42.7 &   \nodata \\
 5858.790 &      26.0 &      4.22 &    --2.260 &   \nodata &      11.4 &   \nodata &      16.9 &   \nodata \\
 5909.970 &      26.0 &      3.21 &    --2.640 &      56.0 &      72.6 &      18.5 &      63.3 &   \nodata \\
 5916.250 &      26.0 &      2.45 &    --2.994 &      92.9 &     105.1 &      51.2 &      95.4 &      17.7 \\
 5927.800 &      26.0 &      4.65 &    --1.090 &      32.4 &      36.7 &      12.2 &      50.9 &   \nodata \\
 5956.690 &      26.0 &      0.86 &    --4.608 &     121.7 &     143.8 &      70.3 &     107.8 &      25.5 \\
 6012.210 &      26.0 &      2.22 &    --4.070 &      44.4 &      58.8 &      11.5 &      53.0 &   \nodata \\
 6019.360 &      26.0 &      3.57 &    --3.360 &       7.0 &   \nodata &   \nodata &      11.3 &   \nodata \\
 6027.050 &      26.0 &      4.07 &    --1.106 &      66.2 &      77.0 &      36.2 &      77.2 &      23.5 \\
 6054.080 &      26.0 &      4.37 &    --2.310 &       5.6 &       8.1 &   \nodata &      12.9 &   \nodata \\
 6105.130 &      26.0 &      4.55 &    --2.050 &   \nodata &      10.9 &   \nodata &      14.8 &   \nodata \\
 6120.240 &      26.0 &      0.91 &    --5.970 &      35.2 &      52.7 &       8.5 &      32.4 &   \nodata \\
 6151.620 &      26.0 &      2.17 &    --3.299 &      88.9 &     102.9 &      46.9 &      87.0 &      18.2 \\
 6157.730 &      26.0 &      4.08 &    --1.320 &   \nodata &      81.2 &      33.2 &      78.1 &      18.8 \\
 6159.380 &      26.0 &      4.61 &    --1.970 &   \nodata &       9.6 &   \nodata &      13.6 &   \nodata \\
 6165.360 &      26.0 &      4.14 &    --1.490 &      40.6 &      51.6 &      14.2 &      54.0 &      10.8 \\
 6173.340 &      26.0 &      2.22 &    --2.880 &     109.6 &     125.6 &      70.7 &     105.6 &      34.9 \\
 6180.200 &      26.0 &      2.73 &    --2.637 &      86.2 &     100.6 &      44.3 &      91.5 &      20.8 \\
 6200.310 &      26.0 &      2.61 &    --2.437 &     105.2 &     122.2 &      70.8 &     105.4 &      36.6 \\
 6219.280 &      26.0 &      2.20 &    --2.433 &     134.3 &     153.3 &     103.3 &     127.4 &      56.7 \\
 6229.230 &      26.0 &      2.84 &    --2.846 &      60.1 &      73.1 &      21.0 &      65.8 &      13.6 \\
 6232.640 &      26.0 &      3.65 &    --1.283 &      92.6 &     104.3 &      58.3 &     103.0 &      35.6 \\
 6246.320 &      26.0 &      3.60 &    --0.894 &     114.0 &     129.2 &      90.5 &     124.1 &      61.7 \\
 6265.130 &      26.0 &      2.17 &    --2.550 &     132.9 &     155.0 &      94.8 &     129.7 &      55.5 \\
 6270.220 &      26.0 &      2.86 &    --2.500 &      76.7 &      93.6 &      41.2 &      79.1 &      17.4 \\
 6271.280 &      26.0 &      3.33 &    --2.728 &      31.6 &      44.9 &      13.4 &      43.6 &   \nodata \\
 6297.790 &      26.0 &      2.22 &    --2.740 &     119.3 &   \nodata &      81.3 &     111.3 &      42.3 \\
 6301.500 &      26.0 &      3.65 &    --0.766 &     115.0 &     130.6 &      91.1 &     117.9 &   \nodata \\
 6322.690 &      26.0 &      2.59 &    --2.426 &     111.0 &     124.4 &      74.2 &     103.5 &      41.6 \\
 6336.820 &      26.0 &      3.68 &    --0.916 &     106.6 &     116.8 &      81.2 &     113.6 &      52.8 \\
 6353.840 &      26.0 &      0.91 &    --6.477 &      11.8 &      23.7 &   \nodata &      14.1 &   \nodata \\
 6355.030 &      26.0 &      2.84 &    --2.403 &   \nodata &     122.2 &      58.9 &   \nodata &      27.8 \\
 6411.650 &      26.0 &      3.65 &    --0.734 &     122.5 &     135.1 &      97.8 &     131.8 &      68.2 \\
 6469.190 &      26.0 &      4.84 &    --0.770 &   \nodata &   \nodata &      18.7 &   \nodata &   \nodata \\
 6574.230 &      26.0 &      0.99 &    --5.004 &      84.1 &     118.5 &      35.4 &      87.0 &   \nodata \\
 6575.020 &      26.0 &      2.59 &    --2.727 &     100.7 &     118.2 &      55.5 &     103.4 &      23.4 \\
 6581.210 &      26.0 &      1.50 &    --4.705 &      72.4 &      87.5 &      21.7 &      69.6 &   \nodata \\
 6609.110 &      26.0 &      2.56 &    --2.692 &     102.0 &     118.8 &      60.9 &     103.1 &      29.1 \\
 6648.080 &      26.0 &      1.01 &    --5.918 &      37.1 &      54.9 &   \nodata &      37.9 &   \nodata \\
 6739.520 &      26.0 &      1.56 &    --4.820 &      43.1 &      60.8 &   \nodata &      44.5 &   \nodata \\
 6750.150 &      26.0 &      2.42 &    --2.621 &     110.9 &     128.9 &      78.6 &     109.5 &      40.2 \\
 6786.860 &      26.0 &      4.19 &    --1.850 &      22.5 &      29.1 &   \nodata &      32.3 &   \nodata \\
 6810.260 &      26.0 &      4.60 &    --1.000 &      36.2 &      46.9 &      14.5 &      54.9 &       8.5 \\
 6971.930 &      26.0 &      3.02 &    --3.390 &      21.0 &      30.3 &   \nodata &      28.7 &   \nodata \\
 7112.170 &      26.0 &      2.99 &    --3.044 &      45.4 &      59.3 &      13.3 &      53.0 &   \nodata \\
 7189.150 &      26.0 &      3.07 &    --2.796 &      57.5 &      66.3 &   \nodata &   \nodata &   \nodata \\
 7223.660 &      26.0 &      3.01 &    --2.269 &      87.2 &     114.4 &      47.2 &      90.8 &   \nodata \\
 7401.690 &      26.0 &      4.18 &    --1.660 &      38.9 &      50.5 &      14.7 &      59.7 &   \nodata \\
 7710.360 &      26.0 &      4.22 &    --1.129 &      62.5 &      76.7 &      32.5 &      75.4 &   \nodata \\
 7723.200 &      26.0 &      2.28 &    --3.617 &      82.1 &     100.4 &      36.6 &      91.9 &   \nodata \\
 7941.090 &      26.0 &      3.27 &    --2.331 &      50.3 &   \nodata &      22.4 &      61.7 &   \nodata \\
 5234.620 &      26.1 &      3.22 &    --2.150 &   \nodata &      81.2 &   \nodata &   \nodata &      58.3 \\
 5325.550 &      26.1 &      3.22 &    --3.220 &   \nodata &      32.1 &      43.4 &   \nodata &      11.6 \\
 5414.070 &      26.1 &      3.22 &    --3.750 &   \nodata &   \nodata &      23.6 &      21.0 &       6.8 \\
 5425.260 &      26.1 &      3.20 &    --3.370 &      27.1 &      38.2 &      41.3 &      37.3 &      14.5 \\
 5991.380 &      26.1 &      3.15 &    --3.557 &      21.5 &   \nodata &      28.3 &      25.8 &       7.7 \\
 6084.110 &      26.1 &      3.20 &    --3.808 &      11.8 &      16.0 &      20.0 &      18.8 &   \nodata \\
 6149.260 &      26.1 &      3.89 &    --2.724 &      21.7 &      25.2 &      30.8 &      26.3 &      11.8 \\
 6247.560 &      26.1 &      3.89 &    --2.329 &      31.2 &      36.7 &      50.2 &      40.3 &      21.4 \\
 6369.460 &      26.1 &      2.89 &    --4.250 &      13.5 &      16.3 &      18.1 &      16.7 &       5.9 \\
 6416.920 &      26.1 &      3.89 &    --2.740 &      23.1 &      29.2 &      29.2 &      31.9 &   \nodata \\
 6432.680 &      26.1 &      2.89 &    --3.708 &      30.0 &      33.6 &      41.7 &      37.5 &      12.9 \\
 6456.380 &      26.1 &      3.90 &    --2.075 &      40.9 &      47.9 &      61.2 &      49.3 &      30.9 \\
 7224.490 &      26.1 &      3.89 &    --3.243 &   \nodata &   \nodata &      12.7 &      13.2 &   \nodata \\
 7711.720 &      26.1 &      3.90 &    --2.543 &      23.7 &      29.1 &      37.4 &      29.8 &      14.9 \\
 6189.000 &      27.0 &      1.71 &    --2.450 &      51.1 &      77.5 &       9.1 &      58.7 &   \nodata \\
 6455.030 &      27.0 &      3.63 &    --0.250 &      19.1 &      29.8 &   \nodata &      31.4 &   \nodata \\
 6632.450 &      27.0 &      2.28 &    --2.000 &      29.1 &      42.3 &   \nodata &      34.7 &   \nodata \\
 5846.990 &      28.0 &      1.68 &    --3.210 &      60.7 &      77.0 &      17.8 &      62.7 &   \nodata \\
 6086.280 &      28.0 &      4.26 &    --0.515 &      31.8 &      36.1 &      12.4 &      47.8 &   \nodata \\
 6175.370 &      28.0 &      4.09 &    --0.535 &      41.4 &      48.6 &      21.1 &      52.5 &      10.4 \\
 6177.240 &      28.0 &      1.83 &    --3.510 &      39.3 &      48.8 &      11.1 &      39.4 &   \nodata \\
 6204.600 &      28.0 &      4.09 &    --1.140 &      14.9 &      18.5 &   \nodata &      25.3 &   \nodata \\
 6635.120 &      28.0 &      4.42 &    --0.828 &      14.9 &      18.2 &   \nodata &      26.0 &   \nodata \\
 6772.320 &      28.0 &      3.66 &    --0.987 &      53.5 &      59.1 &      26.5 &      62.0 &      13.0 \\
 5105.540 &      29.0 &      1.39 &    --1.520 &     139.4 &     160.4 &      67.1 &     123.3 &      32.5 \\
 4722.150 &      30.0 &      4.03 &    --0.390 &      74.1 &      67.9 &      73.6 &      76.5 &      37.4 \\
 4810.528 &      30.0 &      4.08 &    --0.137 &      61.5 &      63.6 &      72.5 &      82.3 &      42.1 \\
 7800.290 &      37.0 &      0.00 &      0.130 &   \nodata &   \nodata &   \nodata &      60.9 &   \nodata \\
 5123.220 &      39.1 &      0.99 &    --0.830 &   \nodata &   \nodata &   \nodata &   \nodata &      16.5 \\
 5200.420 &      39.1 &      0.99 &    --0.570 &     100.9 &   \nodata &      65.8 &      78.6 &      26.5 \\
 5509.910 &      39.1 &      0.99 &    --1.010 &     103.0 &      94.4 &      47.0 &      95.4 &      15.7 \\
 5544.610 &      39.1 &      1.74 &    --1.080 &      29.9 &      18.8 &       9.5 &      16.2 &   \nodata \\
 6127.440 &      40.0 &      0.15 &    --1.060 &      56.3 &      44.9 &   \nodata &      29.7 &   \nodata \\
 6134.550 &      40.0 &      0.00 &    --1.280 &      49.6 &      38.7 &   \nodata &      23.6 &   \nodata \\
 6143.200 &      40.0 &      0.07 &    --1.100 &      54.4 &      44.6 &   \nodata &      30.7 &   \nodata \\
 5112.270 &      40.1 &      1.67 &    --0.590 &   \nodata &   \nodata &      21.8 &   \nodata &   \nodata \\
 5853.640 &      56.1 &      0.60 &    --1.010 &     143.5 &     121.4 &     110.8 &      99.8 &      49.6 \\
 6141.730 &      56.1 &      0.70 &    --0.080 &     213.6 &     177.6 &     166.4 &     151.1 &      93.1 \\
 5805.770 &      57.1 &      0.13 &    --1.560 &      47.7 &      28.8 &      13.0 &      25.1 &   \nodata \\
 6390.480 &      57.1 &      0.32 &    --1.410 &      43.0 &      22.0 &      11.6 &      18.6 &   \nodata \\
 5274.240 &      58.1 &      1.04 &    --0.320 &      29.1 &   \nodata &      12.7 &   \nodata &   \nodata \\
 5472.300 &      58.1 &      1.24 &    --0.180 &      19.4 &   \nodata &   \nodata &   \nodata &   \nodata \\
 4959.120 &      60.1 &      0.06 &    --0.800 &      79.9 &      74.0 &      35.0 &      64.2 &   \nodata \\
 5092.790 &      60.1 &      0.38 &    --0.610 &      47.3 &   \nodata &      21.9 &      20.9 &   \nodata \\
 5212.360 &      60.1 &      0.20 &    --0.960 &      61.4 &      61.9 &      19.6 &      51.6 &   \nodata \\
 5234.190 &      60.1 &      0.55 &    --0.510 &      71.9 &      57.9 &      17.9 &      47.3 &   \nodata \\
 5249.580 &      60.1 &      0.98 &      0.200 &      64.5 &      54.7 &      31.8 &      50.7 &   \nodata \\
 5255.510 &      60.1 &      0.20 &    --0.670 &     116.3 &   \nodata &      40.5 &   \nodata &   \nodata \\
 5293.160 &      60.1 &      0.82 &      0.100 &      74.2 &      71.6 &      29.0 &      62.5 &   \nodata \\
 5319.810 &      60.1 &      0.55 &    --0.140 &      86.6 &   \nodata &      34.4 &   \nodata &   \nodata \\
 6645.110 &      63.1 &      1.38 &      0.204 &      24.5 &      27.8 &      17.0 &      30.3 &   \nodata \\
\enddata
\end{deluxetable*}

\subsection{Model Atmospheres}
We have employed the ATLAS9 plane-parallel stellar atmospheres of \citet{castelli;kurucz_2003}. These one-dimensional models ignore any centre-to-limb spatial variations, assume hydrostatic equilibrium and no convective overshoot from the photosphere. The stellar parameter spacing between models is 250\,K, 0.5\,dex in surface gravity, 0.5\,dex in [M/H] and 0.1\,dex in [$\alpha$/Fe]. We interpolated the temperature, gas and radiative pressure, electron density and opacities between atmosphere models using the Quickhull algorithm \citep{barber;et-al_1996}. Quickhull is reliant on Delaunay tessellation, which suffers from extremely skewed cells when the grid points vary in size by orders of magnitude -- as $T_{\rm eff}$ values do compared to $\log{g}$ or [X/H]. If unaccounted for, interpolating using such asymmetric cells can result in significant errors in atmospheric properties across all photospheric depths. We scaled each stellar parameter between zero and unity prior to interpolation in order to minimise these effects. During this step, $\log{g}$, [M/H] and [$\alpha$/Fe] values were scaled such that we simultaneously interpolate linearly in $T_{\rm eff}$ and logarithmically in all other parameters.

\begin{deluxetable*}{lcccccccccccccc}[h!]
\tablecolumns{2}
\tabletypesize{\scriptsize}
\tablecaption{Stellar Parameters\label{tab:stellar-parameters}}
\tablehead{
	\colhead{Designation} &
	\colhead{$T_{\rm eff}$} &
	\colhead{$\log{g}$} &
	\colhead{$v_t$} &
	\colhead{[Fe/H]} & 
	\colhead{$T_{\rm eff}$} &
	\colhead{$\log{g}$} &
	\colhead{$v_t$} &
	\colhead{[Fe/H]} &
	\colhead{Source} \\
	& (K) & (dex) & (km\,s$^{-1}$) & (dex) & (K) & (dex) & (km s$^{-1}$) & (dex) 
}
\startdata
\multicolumn{10}{c}{Standard Stars} \\
\cline{1-10} \\
HD41667			& 4630	& 1.70	& 1.66 	& --1.21
				& 4605	& 1.88	& 1.44	& --1.16
				& \citet{gratton;et-al_2000} \\
HD44007			& 4790	& 1.78	& 1.63	& --1.80
				& 4850	& 2.00 	& 2.20	& --1.71
				& \citet{fulbright_2000} \\
HD142948		& 4950	& 2.19	& 1.78	& --0.79
				& 4713 	& 2.17 	& 1.38	& --0.77
				& \citet{gratton;et-al_2000} \\
\\
\multicolumn{10}{c}{Program Stars} \\
\cline{1-10} \\
C2225316-14437	& 4325				& 1.26				& 1.78				& --1.26	
				& 4235 $\pm$ 118 	& 1.45 $\pm$ 0.21	& 1.96 $\pm$ 0.11 	& --1.20 $\pm$ 0.14 
				& \citet{wylie-de-boer;et-al_2012} \\
C2306265-085103	& 4170	& 0.85	& 1.81 	& --1.17
				& \dots	& \dots	& \dots	& \dots
				& \dots \\	
J221821-183424	& 4590				& 0.90				& 1.91				& --1.61
				& 4395 $\pm$ 205 	& 1.45 $\pm$ 0.35 	& 1.96 $\pm$ 0.18 	& --1.15 $\pm$ 0.21
				& \citet{wylie-de-boer;et-al_2012} \\
J223504-152834	& 4590				& 2.15 				& 1.45				& --0.68
				& 4597 $\pm$ 158 	& 2.40 $\pm$ 0.14 	& 1.47 $\pm$ 0.07 	& --0.98 $\pm$ 0.17
				& \citet{wylie-de-boer;et-al_2012} \\
J223811-104126	& 5140				& 2.96				& 1.33				& --1.45
				& 5646 $\pm$ 147 	& 4.60 $\pm$ 0.15 	& 1.09 $\pm$ 0.11 	& --1.20 $\pm$ 0.20
				& \citet{wylie-de-boer;et-al_2012} 
\enddata

\end{deluxetable*}

\subsection{Stellar Parameters}
The most recent version of the spectral synthesis code \textsc{moog} \citep{sneden;et-al_1973} has been used to derive individual line abundances and stellar parameters. This version employs Rayleigh scattering \citep{sobeck;et-al_2011} instead of treating scattering as true absorption, which is particularly important for transitions blue-ward of 450\,nm. This is noteworthy, but is less relevant for these analyses as most of our line measurements are red-ward of 450\,nm. Absorption lines are assumed to form under the assumption of Local Thermal Equilibrium (LTE).

\begin{figure*}[t!]
	\includegraphics[width=\textwidth]{./figures/spectra-h-alpha.pdf}
	\caption{Sample spectra of the H-$\alpha$ region for all Aquarius stream candidates.}
	\label{fig:spectra-h-alpha}
\end{figure*}

\subsubsection{Effective Temperature}
\label{sec:effective-teffs}
The effective temperature for each star was found by demanding a zero-trend in excitation potential and line abundance for each measured Fe\,\textsc{i} transition. In the same way, the microturbulence was found by demanding a zero trend in reduced equivalent width against abundance. Linear relationships with slopes of $|0.001|$ dex were considered to be converged.

\subsubsection{Photometric Effective Temperatures}
\label{sec:photometric-temperatures}

The effective temperatures published for all of our stars in Table \ref{sec:stellar-parameters} have been derived spectroscopically. As a consistency check for our spectroscopic temperatures, we have estimated effective temperatures using infrared photometry from the 2MASS catalog and empirical $T_{\rm eff}$-colour relationships. The $J-K$ color has been employed for these calculations because it has the least dependence on metallicity.

The Aquarius stream is relatively close and at high-latitude, so significant extinction caused by dust is not expected. The reddening maps of \citet{schlegel;et-al_1998} estimate that the extinction for our stars varies between E(B-V) = 0.03 to 0.07\,mags. However, these maps are not perfect. As an alternative to using the \citet{schlegel;et-al_1998} dust maps, the level of extinction towards a star can be found by the strength of the interstellar Na D or K \textsc{i} lines \citep{munari;zwitter_1997}. The interstellar lines must be sufficiently separated from the stellar absorption lines in order for the extinction to be measurable. This was the case in only one of our stars, J223811. The equivalent width of the Na D1 line in J223811 is found to be 302.4\,m{\AA}, corresponding to an $\mbox{E(B-V)} = 0.12^{+0.01}_{-0.02}$ according to the \citet{munari;zwitter_1997} relationship. This is a significantly higher value than the \citet{schlegel;et-al_1998} reddening map suggests: $\mbox{E(B-V)} = 0.07$. For completeness, both extinction values have been used to calculate photometric temperatures.

Although all of our candidates are giant stars, two different $T_{\rm eff}$-colour relationships have been employed: the \citet{alonso;et-al_1999} calibration for giants, and the \citet{casagrande;et-al_2010} calibration for dwarfs and sub-giants. All photometrically-derived temperatures are listed in \ref{tab:photometric-temperatures}. The \citet{alonso;et-al_1999} $J-K$ scale is insensitive to metallicity effects, and is in good agreement with our spectroscopic temperatures. For the \citet{casagrande;et-al_2010} $J-K$ relationship, which has a slight metallicity dependency, we have utilized the spectroscopically-derived [Fe/H] values shown in Table \ref{tab:stellar-parameters}.

The effective temperatures found from spectroscopy agree fairly well with the photometric estimates. For three of our stars, the temperatures we find are in between the two color-temperature relationships. There are two noteworthy deviations: J223504-152834 and J223811-104126. For J223504-152834 we find the star to have an effective temperature of 4590\,K, only 10\,K hotter than the \citet{casagrande;et-al_2010} dwarf/sub-giant relationship, yet ${\sim}$500\,K hotter than the \citet{alonso;et-al_1999} estimate. Our temperature of 4590\,K is in good agreement with previous spectroscopic determinations (4795\,K; \citet{williams;et-al_2011}, 4597\,K; \citet{wylie-de-boer;et-al_2012}). The fact that the \citet{alonso;et-al_1999} photometric temperature relationship for giants predicts a temperature that is ${\sim}$500\,K cooler than the spectroscopically-derived values may be attributed to an underestimated amount of dust towards J223504-152834. A higher than expected level of extinction along the line-of-sight would produce cooler photometric temperatures than the true effective temperature.  An extinction of $E(B-V) = 0.10$ would be required for the \citet{alonso;et-al_1999} relation to yield a temperature of ${\sim}$4500\,K \--- within the uncertainties of our effective temperature derived from spectroscopy. Such a discrepancy in extinction is large, but not uncommon \citep[e.g. see][]{peek;graves_2010,nataf;et-al_2012}. Although we suspect J223504-152834 to have a higher reddening than $E(B-V) = 0.04$, we have adopted the \citet{schlegel;et-al_1998} value for this analysis.

J223811-104126, the second temperature deviant, has two extinction values available: the \citet{schlegel;et-al_1998} value of $E(B-V) = 0.07$ and the 0.12 value determined from the interstellar Na D1 line. Given the inherent difficulties in producing dust maps, we believe the $E(B-V) = 0.12$ determination to be more accurate. However, using the \citet{schlegel;et-al_1998} extinction value with the \citet{alonso;et-al_1999} color-temperature relationship yields excellent agreement (5104\,K) with our spectroscopic temperature of 5140\,K. Employing our more accurate reddening determination yields 5215\,K, still within the uncertainties of the relationship. The \citet{casagrande;et-al_2010} scale is ${\sim}$250\,K hotter in both cases.

With the exception of J223504-152834, the \citet{alonso;et-al_1999} giant relationship agrees well ($\pm$100\,K) with our spectroscopically-derived values. The \citet{casagrande;et-al_2010} relationship is also in relatively good agreement (most within 100\,K, largest discrepancy is 220\,K) with our spectroscopic temperatures, although we note that this relation was derived primarily for dwarfs/sub-giants and our candidates are all giant stars.

\begin{deluxetable}{lccccc}
\tablecolumns{1}
\tabletypesize{\scriptsize}
\tablecaption{Reddening \& Photometric Temperatures\label{tab:photometric-temperatures}}
\tablehead{
& & & \multicolumn{2}{c}{$T_{phot}$} \\
\cline{4-5} \\
	\colhead{Designation} &
	\colhead{E($B-V$)} &
	\colhead{$(J-K)_0$} &
	\colhead{Alonso} &
	\colhead{Casagrande} \\
	& (mag) & & (K) & (K)
}
\startdata 
C222531-144370 	& 0.03\phn	& 	0.75 & 4247	&	4391 \\	
C230626-085103	& 0.05\phn	&	0.82 & 4091	&	4200 \\	
J221821-183424 	& 0.03\phn	&	0.64 & 4546	&	4747 \\	
J223504-152834 	& 0.04\phn	&	0.69 & 4110	& 	4579 \\	
J223811-104126 	& 0.07\phn	&	0.48 & 5104	&	5367 \\	
				& 0.12\footnote{Extinction derived from interstellar Na D1 line (see \ref{sec:photometric-temperatures})} 			&	0.45 & 5215	&	5484	
\enddata
\end{deluxetable}

\subsubsection{Surface Gravity}

The surface gravity for all stars was found by forcing the mean Fe\,\textsc{i} and Fe\,\textsc{ii} abundances to be equal. The process is iterative: a zero trend with the excitation potential, reduced equivalent width and abundance must be maintained. For two stars, X and Y, the surface gravity could not be found by balancing the Fe\,\textsc{i} and \textsc{ii} line abundances. The smallest difference between the median abundances was $\langle$Fe\,\textsc{i} $-$ Fe\,\textsc{ii}$\rangle = 0.05$\,dex, which was considered to be converged. Varying the surface gravity from this solution did alter the mean Fe\,\textsc{i} and \textsc{ii} abundances, but the difference $\langle$Fe\,\textsc{i} $-$ Fe\,\textsc{ii}$\rangle$ remained constant at 0.05\,dex. In these two instances, the surface gravity was set to be closer towards the isochrone, while the $\langle$Fe\,\textsc{i} $-$ Fe\,\textsc{ii}$\rangle$ = 0.05\,dex condition was maintained.

\subsubsection{Metallicity, $\mbox{[M/H]}$}
The degeneracy between parameters required a refinement of the stellar parameters until the line abundances were consistent with the model atmosphere. After these parameters had converged, the model metallicity was exactly matched to that of our mean Fe\,\textsc{i} abundance. Line abundances that were unusually deviant ($>$3-$\sigma$) from the mean were removed. The largest number of outlier measurements removed for any observation was nine for C222531-144370, leaving 60 Fe\,\textsc{i} and 10 Fe\,\textsc{ii} lines for analysis. The minimum number of Fe transitions used in any star was 44 lines (33 Fe\,\textsc{i} and 11 Fe\,\textsc{ii}) for our hottest star, J223811-104126.

% spectral plots for five stars

\subsection{Abundances}
We have scaled our abundances to solar values using the Sun's chemical composition described in \citet{asplund;et-al_2009}.

\subsubsection{Carbon}

\subsubsection{Oxygen}
\label{sec:oxygen-abundances}
Oxygen is a particularly difficult element to measure. There are only a handful of lines available in an optical spectrum: the forbidden [O \textsc{i}] lines at 630 nm and 636 nm, and the O \textsc{i} triplet lines at 777 nm. The forbidden lines are very weak, and become immeasurable in either hot and/or metal-poor stars ([Fe/H] $\lesssim -1.5$ dex). When they are present, depending on the radial velocity of the star, the [O \textsc{i}] lines can be significantly affected by telluric absorption. Moreover, the 630 nm line is blended with a Ni \textsc{i} absorption line \citep{allende-prieto;et-al_2001}, and hence the region requires careful consideration. Although the O \textsc{i} triplet lines at 777 nm are stronger than the forbidden lines, they are extremely susceptible to non-LTE and 3D effects \citep{asplund;perez_2001}, and sensitive to changes in microturbulence.

The [O \textsc{i}] lines were measurable in four of our candidates. Careful consideration was taken when correcting for telluric absorption. The 6300$\lambda$ line in one of our candidates, C2306265-085103, was sufficiently affected by telluric absorption such that the line was considered immeasurable. Thus, only the 6363$\lambda$ transition was used to derive an oxygen abundance for C2306265-085103. In our hottest star, J223811-104126, the forbidden oxygen lines were not detected above a 3-$\sigma$ significance. After synthesising the region, we deduce a very conservative upper limit of $\mbox{[O/Fe]} < -0.50$ dex from the [O \textsc{i}] lines. This is consistent with the rest of our candidates, with [O/Fe] abundances varying between 0.41 to 0.57\,dex.

In order to derive an oxygen measurement for J223811-104126, we were forced to use the triplet lines at 777\,nm. Each line was synthesised, and a mean abundance was calculated for each candidate. Abundances determined from the oxygen triplet lines were systematically $+0.27$\,dex higher than found from the [O \textsc{i}] lines. \citet{perez;et-al_2006} found [O/Fe] values based on the O \textsc{i} permitted triplet lines are on average $+0.19 \pm 0.07$\,dex higher than those found from the forbidden lines, after non-LTE corrections of $-$0.08\,dex had been applied to the permitted abundance.  Thus, our 0.27\,dex offset between measurements of the [O \textsc{i}] and O \textsc{i} triplet lines is exactly the same as found by \citet{perez;et-al_2006}, which we attribute to a combination of systematic phenomena including non-LTE and 3D effects. \citet{perez;et-al_2006} concluded the weak forbidden lines, when not too weak, probably give the most reliable estimate of oxygen abundance.

From the permitted O \textsc{i} triplet, we derive an oxygen abundance of $\mbox{[O/Fe]} = 0.41 \pm 0.01$\,dex (random scatter) for J223811-104126. This measurement will be systematically higher than the `true' abundance if it were discernible from the [O \textsc{i}] lines, on the order of ${\sim}$0.27\,dex. When we apply this crude offset derived from the rest of our sample, we arrive at a corrected abundance of $\mbox{[O/Fe]} = 0.14 \pm 0.08$\,dex for J223811-104126. This is the most oxygen-deficient star in our sample.

\subsubsection{Sodium}
\label{sec:sodium-abundances}
Our line list includes three clean, unblended sodium lines at 5688$\lambda$, 6154$\lambda$ and 6161$\lambda$. Not all three of these lines were detectable in each star. In the hottest and most metal-poor stars, J223811 and J221821-183424 respectively, only the 5688$\lambda$ line was measurable. Consequently, an uncertainty of 0.10\,dex was adopted for these stars. For stars where multiple sodium lines were available, the line-to-line scatter is reasonable: smaller than 0.09\,dex.

% [Na/Fe] measurements in the standard stars?

\subsubsection{$\alpha$-elements (Mg, Ca, Si and Ti)}
\label{sec:alpha-elements}

The $\alpha$-elements (Mg, Ca, Si and Ti) are forged through $\alpha$-particle capture in assorted burning stages of stellar evolution, including the burning of carbon, neon and silicon. Although Ti ($Z = 22$) is not formally an $\alpha$-element, its abundance generally varies with those of the other $\alpha$-elements and therefore has been included here.

% magnesium measurements
%Magnesium has been measured from unblended lines that fall within the linear regime between reduced equivalent width and abundance.
Depending on the radial velocity of the star, some magnesium lines were be affected by telluric absorption, particularly the 6318$\lambda$ and 6965$\lambda$ transitions. Atmospheric absorption was most notable for C2225316-14437, where three of the four Mg transitions in our line list suffered some degree of telluric absorption, requiring an attentive correction. Every amended absorption profile was carefully examined, and lines with suspicious profiles were excluded from the final magnesium abundance.

% [alpha/Fe]
\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./figures/aquarius-alpha-fe.pdf}
	\caption{A plot showing $[\alpha/\mbox{Fe}]$ for all Aquarius stream stars.}
	\label{fig:alpha-fe}
\end{figure}

% calcium measurements + silicon measurements
Of all the $\alpha$-elements, calcium has the smallest measurement scatter in our stars. Four line measurements were used for each star, with a measurement uncertainty of 0.01\,dex. As shown in Figure \ref{fig:alpha-elements}, all Aquarius candidates show super-solar Ca abundances, ranging between 0.23 to 0.43 dex, consistent with both [Mg/Fe] and [Si/Fe] measurements.

However, C2225316-14437 has an unusually high silicon abundance ($\mbox{[Si/Fe]} = 0.72$), well outside the uncertainties of the rest of our sample. The silicon line abundances are in relatively good agreement with one another. If we exclude the most prominent outlier, then the mean abundance drops only slightly to $\mbox{[Si/Fe]} = 0.70 \pm 0.04$\,dex. The lowest silicon line abundance for C2225316-14437 is $\mbox{[Si/Fe]} = 0.57$\,dex, which is still significantly higher than the averaged abundance for any other star.

% titanium measurements?
Titanium abundances for the stream stars are generally in agreement, showing typical $\alpha$-enhancement signatures. Our titanium measurements are derived from clean, unblended lines. The number of transitions available in our spectra varies from four to seven. In our hottest and most metal-poor stars, no suitable Ti\,\textsc{i} transitions were available. In other stream candidates, the mean difference between species is $\langle$Ti\,\textsc{i}$ - $Ti\,\textsc{ii}$\rangle = 0.24$\,dex. Although this discrepancy is not negligible, there are limited lines available for comparison, and our $\langle$Fe\,\textsc{i}$ - $Fe\,\textsc{ii}$\rangle$ abundances are in good agreement.

\subsubsection{Aluminium}
There are six aluminium transition lines available in our optical spectra. The strongest of these lines occur at 3944 and 3961\,{\AA}, and are visible in all of our stars. However, this is a particularly crowded spectral region: the lines fall between the strong Ca H and K lines, with the 3961\,{\AA} transition clearly in the wing of the Ca H line. This makes the continuum determination and subsequent aluminium measurement non-trivial, especially given the metallicity of our stars. Additionally, the 3944 and 3961 lines suffer from appreciable departures from LTE, resulting in under-estimated abundances by up to $\sim$0.60\,dex \citep{baumueller;gehren_1997}. Instead, we have measured Al abundances from other available transitions: the Al \textsc{i} lines at 6696, 6698, 7835 and 7836\,{\AA}. Although these lines are weaker than the bluer Al \textsc{i} lines, their surrounding regions are uncrowded from strong absorption lines.

Generally the four Al \textsc{i} lines are in reasonable agreement with one another, yielding random scatter of less than 0.05\,dex. However, these lines were not always detected in the data. The strength of the lines is noticeably weakened below a 3-$\sigma$ detection with decreasing metallicity and increasing temperature. For our most metal-poor star, J221821, only a single Al \textsc{i} line was detectable, and so we have adopted a conservative uncertainty of 0.10\,dex.

\subsubsection{Iron-peak Elements (Sc, V, Cr and Mn)}

As expected, the Fe-peak elemental abundances exhibit similar trends (Figure \ref{fig:fe-peak-elements}). The noteworthy exception is chromium, where the median abundance is $\mbox{[Cr/Fe]} = -0.08 \pm 0.12$ and does not appear to vary with overall metallicity. The number of clean, suitable Cr \textsc{i} lines available between candidates fluxuated from three to twelve. Very little line-to-line scatter is present in both Cr \textsc{i} and \textsc{ii}: the random scatter is below 0.04\,dex for most stars, with the isolated example of J223504 where $\sigma(\mbox{[Cr \textsc{ii}/Fe]}) = 0.29$\,dex, and only three lines were available. The Cr \textsc{i} and \textsc{ii} abundances for J223504 are in good agreement (0.05\,dex), and since none of the measured profiles were particularly unusual, we chose not to exclude outlier measurements. For all other stars, the Cr \textsc{i} $-$ Cr \textsc{ii} abundance differences varied between 0.02\,dex up to 0.31\,dex. The largest discrepancy occurred in our hottest star, where fewer chromium transitions were detected.

All other iron peak elements considered (Sc, V, Mn) all exhibit a slight downward trend in metallicity, with manganese demonstrating the most prominent gradient: varying between $\mbox{[Mn/Fe]} = -0.52$\,dex to $+0.31$\,dex in the most metal-poor and -rich stars respectively. The [Sc/Fe] measurements presented in Figure \ref{fig:fe-peak-elements} are calculated from six clean Sc \textsc{ii} lines, and there is very minimal line-to-line scatter, the largest of which is 0.06\,dex. Three Aquarius stream stars display [Sc/Fe] $\sim$ +0.20, while J221821 and J223811 have $\mbox{[Sc/Fe]} = 0.01$ and 0.05\,dex, respectively.

% Iron peak elements
\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./figures/aquarius-fe-peak.pdf}
	\caption{Iron peak elements for all stars.}
	\label{fig:fe-peak-elements}
\end{figure}

\subsubsection{Nickel}
Most Aquarius stream stars have seven clean Ni\,\textsc{i} transitions available. These lines are in excellent agreement, with typical scatter of 0.03\,dex.
% nickel abundances in standard stars

\subsubsection{Barium}
%Barium is a neutron-capture element, primarily produced through the \textit{s}-process, but can be forged from a variety of nucleosynthetic processes.

Measuring barium requires some careful consideration. The strong resonance lines at 4554 {\AA} and 4934 {\AA} have appreciable hyperfine splitting, which can result in over-estimated abundances if unaccounted for. There are a few singly ionized lines available in our spectra that do not suffer from strong hyperfine splitting, namely the 5853, 6141 and 6496 {\AA} transitions. Depending on the radial velocity of the star, sometimes our 6496 {\AA} line was affected by small amounts of telluric absorption. In these cases the line was discarded. The abundances tabulated in Table \ref{tab:program-star-abundances} are based on the remaining clean Ba \textsc{ii} lines at 5853 and 6141 {\AA}, which are generally in good agreement with each other, and the 6496 {\AA} line where available.

Our standard stars have [Ba/Fe] abundances very typical of the Milky Way halo. Only one of our standard stars, HD 44007, has published [Ba/Fe] abundances from high-resolution spectra. We find HD 44007 to have $\mbox{[Ba/Fe]} = -0.05 \pm 0.04$, which is at the mid-point of existing published measurements from \citet{fulbright_2000} and \citet{burris;et-al_2000} of $-0.27$ and 0.05 dex, respectively.

With one exception, the Aquarius stream candidates have barium abundances that are quite similar to the stellar halo, ranging between $\mbox{[Ba/Fe]} = -0.02$ to 0.15 dex. The exception is C2225316, which has an anomalously high barium abundance of $\mbox{[Ba/Fe]} = 0.70$ dex. This is ${\sim}$0.70 dex higher than the Milky Way trend at its given $\mbox{[Fe/H]} = -1.26$ dex. Our three Ba \textsc{ii} lines in C2225316 are in excellent agreement: 0.67, 0.68, and 0.72 dex. The 6496\,{\AA} transition in C2225316 has a negligible telluric contribution and was included for this candidate. At least in Al, Si, Cr and Ba content (see sections \ref{sec:aluminium-abundance}, \ref{sec:alpha-elements}, \ref{sec:iron-peak-elements}), C225316 appears to be chemically distinct from the other Aquarius stream stars in our sample.

\begin{deluxetable*}{lccccccclcccccc}
\tablecolumns{15}
\tablecaption{Program Star Abundances\label{tab:program-star-abundances}}
\tablehead{
\colhead{Species} & $N$ & $\log\epsilon(X)$ & $\sigma_\epsilon$ & [X/H] & [X/Fe] & $\sigma$ &&
\colhead{Species} & $N$ & $\log\epsilon(X)$ & $\sigma_\epsilon$ & [X/H] & [X/Fe] & $\sigma$ 
}
\startdata
\\
 \multicolumn{7}{c}{J221821-183424} & \colhead{} & \multicolumn{7}{c}{C2225316-14437} \\
 \cline{1-7} \cline{9-15} \\
   O \textsc{i} &   2 &    7.55 &    0.04 & $-$1.13 &    0.47 &    0.02 &&
   O \textsc{i} &   2 &    7.92 &    0.00 & $-$0.77 &    0.49 &    0.00 \\
  Na \textsc{i} &   1 &    4.71 &    0.00 & $-$1.53 &    0.08 &    0.00 &&
  Na \textsc{i} &   2 &    5.07 &    0.02 & $-$1.17 &    0.09 &    0.01 \\
  Mg \textsc{i} &   3 &    6.33 &    0.09 & $-$1.27 &    0.34 &    0.05 &&
  Mg \textsc{i} &   2 &    6.84 &    0.09 & $-$0.76 &    0.50 &    0.06 \\
  Al \textsc{i} &   1 &    5.05 &    0.00 & $-$1.40 &    0.21 &    0.00 &&
  Al \textsc{i} &   4 &    5.87 &    0.09 & $-$0.58 &    0.68 &    0.04 \\
  Si \textsc{i} &   5 &    6.28 &    0.08 & $-$1.23 &    0.38 &    0.04 &&
  Si \textsc{i} &   5 &    6.97 &    0.15 & $-$0.54 &    0.72 &    0.07 \\
  Ca \textsc{i} &   4 &    4.99 &    0.04 & $-$1.35 &    0.25 &    0.02 &&
  Ca \textsc{i} &   4 &    5.51 &    0.04 & $-$0.83 &    0.43 &    0.02 \\
 Sc \textsc{ii} &   6 &    1.56 &    0.09 & $-$1.59 &    0.01 &    0.04 &&
 Sc \textsc{ii} &   5 &    2.08 &    0.13 & $-$1.07 &    0.19 &    0.06 \\
  Ti \textsc{i} &   0 & \nodata & \nodata & \nodata & \nodata & \nodata &&
  Ti \textsc{i} &   4 &    4.04 &    0.02 & $-$0.91 &    0.36 &    0.01 \\
 Ti \textsc{ii} &   4 &    3.85 &    0.13 & $-$1.10 &    0.50 &    0.06 &&
 Ti \textsc{ii} &   2 &    4.29 &    0.08 & $-$0.66 &    0.61 &    0.05 \\
   V \textsc{i} &   3 &    2.26 &    0.01 & $-$1.67 & $-$0.06 &    0.01 &&
   V \textsc{i} &   6 &    2.93 &    0.17 & $-$1.00 &    0.26 &    0.07 \\  
  Cr \textsc{i} &  11 &    3.77 &    0.08 & $-$1.87 & $-$0.26 &    0.02 &&
  Cr \textsc{i} &   8 &    4.29 &    0.16 & $-$1.35 & $-$0.08 &    0.06 \\
 Cr \textsc{ii} &   2 &    4.04 &    0.02 & $-$1.60 &    0.00 &    0.02 &&
 Cr \textsc{ii} &   1 &    4.27 &    0.00 & $-$1.37 & $-$0.11 &    0.00 \\
  Mn \textsc{i} &   3 &    3.30 &    0.03 & $-$2.13 & $-$0.52 &    0.02 &&
  Mn \textsc{i} &   3 &    4.15 &    0.04 & $-$1.28 & $-$0.01 &    0.03 \\
  Fe \textsc{i} &  52 &    5.89 &    0.09 & $-$1.61 &    0.00 &    0.01 &&
  Fe \textsc{i} &  60 &    6.24 &    0.10 & $-$1.26 &    0.00 &    0.01 \\
 Fe \textsc{ii} &  13 &    5.91 &    0.05 & $-$1.59 &    0.02 &    0.01 &&
 Fe \textsc{ii} &  10 &    6.19 &    0.06 & $-$1.31 & $-$0.05 &    0.02 \\
  Co \textsc{i} &   1 &    3.38 &    0.00 & $-$1.61 & $-$0.00 &    0.00 &&
  Co \textsc{i} &   3 &    3.92 &    0.12 & $-$1.07 &    0.19 &    0.07 \\
  Ni \textsc{i} &   5 &    4.62 &    0.14 & $-$1.60 &    0.01 &    0.06 &&
  Ni \textsc{i} &   7 &    5.05 &    0.09 & $-$1.17 &    0.09 &    0.03 \\
  Cu \textsc{i} &   1 &    2.16 &    0.00 & $-$2.03 & $-$0.42 &    0.00 &&
  Cu \textsc{i} &   1 &    3.38 &    0.00 & $-$0.81 &    0.45 &    0.00 \\
  Zn \textsc{i} &   2 &    3.17 &    0.11 & $-$1.39 &    0.22 &    0.08 &&
  Zn \textsc{i} &   2 &    3.48 &    0.24 & $-$1.08 &    0.19 &    0.17 \\
 Ba \textsc{ii} &   2 &    0.72 &    0.08 & $-$1.47 &    0.14 &    0.05 &&
 Ba \textsc{ii} &   2 &    1.62 &    0.01 & $-$0.56 &    0.70 &    0.00 \\

\cline{1-7} \cline{9-15} \\
\\
\multicolumn{7}{c}{J223504-152834} && \multicolumn{7}{c}{J223811-104126} \\
\cline{1-7} \cline{9-15} \\
   O \textsc{i} &   2 &    8.46 &    0.10 & $-$0.23 &    0.45 &    0.07 &&
   O \textsc{i} &   1 &    8.26 &    0.00 & $-$0.43 &    0.14\footnote{Abundance derived from the permitted O \textsc{i} triplet instead of the forbidden [O \textsc{i}] lines, see \S\ref{sec:oxygen-abundances}} &    0.08 \\
  Na \textsc{i} &   3 &    5.74 &    0.12 & $-$0.50 &    0.18 &    0.07 &&
  Na \textsc{i} &   1 &    4.86 &    0.00 & $-$1.38 &    0.07 &    0.00 \\
  Mg \textsc{i} &   3 &    7.41 &    0.15 & $-$0.19 &    0.49 &    0.09 &&
  Mg \textsc{i} &   2 &    6.47 &    0.02 & $-$1.13 &    0.32 &    0.01 \\
  Al \textsc{i} &   3 &    6.08 &    0.08 & $-$0.37 &    0.31 &    0.05 &&
  Al \textsc{i} &   2 &    5.10 &    0.14 & $-$1.35 &    0.10 &    0.10 \\
  Si \textsc{i} &   5 &    7.12 &    0.10 & $-$0.39 &    0.29 &    0.04 &&
  Si \textsc{i} &   3 &    6.38 &    0.04 & $-$1.13 &    0.32 &    0.02 \\
  Ca \textsc{i} &   4 &    6.00 &    0.03 & $-$0.34 &    0.33 &    0.02 &&
  Ca \textsc{i} &   4 &    5.30 &    0.03 & $-$1.04 &    0.41 &    0.01 \\
 Sc \textsc{ii} &   6 &    2.70 &    0.14 & $-$0.45 &    0.23 &    0.06 &&
 Sc \textsc{ii} &   6 &    1.75 &    0.14 & $-$1.40 &    0.05 &    0.06 \\
  Ti \textsc{i} &   4 &    4.59 &    0.02 & $-$0.36 &    0.32 &    0.01 &&
  Ti \textsc{i} &   0 & \nodata & \nodata & \nodata & \nodata & \nodata \\
 Ti \textsc{ii} &   2 &    4.91 &    0.27 & $-$0.04 &    0.64 &    0.19 &&
 Ti \textsc{ii} &   4 &    3.83 &    0.09 & $-$1.12 &    0.33 &    0.04 \\
   V \textsc{i} &   6 &    3.62 &    0.22 & $-$0.31 &    0.37 &    0.09 &&
   V \textsc{i} &   1 &    2.42 &    0.00 & $-$1.51 & $-$0.06 &    0.00 \\
  Cr \textsc{i} &   7 &    4.88 &    0.09 & $-$0.76 & $-$0.08 &    0.04 &&
  Cr \textsc{i} &  12 &    4.12 &    0.06 & $-$1.52 & $-$0.07 &    0.02 \\
 Cr \textsc{ii} &   3 &    4.83 &    0.50 & $-$0.81 & $-$0.13 &    0.29 &&
 Cr \textsc{ii} &   3 &    4.34 &    0.07 & $-$1.30 &    0.15 &    0.04 \\
  Mn \textsc{i} &   3 &    5.06 &    0.10 & $-$0.37 &    0.31 &    0.06 &&
  Mn \textsc{i} &   3 &    3.55 &    0.04 & $-$1.88 & $-$0.43 &    0.02 \\
  Fe \textsc{i} &  63 &    6.82 &    0.12 & $-$0.68 &    0.00 &    0.02 &&
  Fe \textsc{i} &  33 &    6.05 &    0.06 & $-$1.45 &    0.00 &    0.01 \\
 Fe \textsc{ii} &  12 &    6.82 &    0.07 & $-$0.68 &    0.00 &    0.02 &&
 Fe \textsc{ii} &  11 &    6.00 &    0.09 & $-$1.50 & $-$0.05 &    0.03 \\
  Co \textsc{i} &   3 &    4.50 &    0.13 & $-$0.49 &    0.19 &    0.08 &&
  Co \textsc{i} &   0 & \nodata & \nodata & \nodata & \nodata & \nodata \\
  Ni \textsc{i} &   7 &    5.61 &    0.09 & $-$0.61 &    0.07 &    0.03 &&
  Ni \textsc{i} &   2 &    4.82 &    0.04 & $-$1.40 &    0.05 &    0.03 \\
  Cu \textsc{i} &   1 &    3.82 &    0.00 & $-$0.37 &    0.31 &    0.00 &&
  Cu \textsc{i} &   1 &    2.39 &    0.00 & $-$1.80 & $-$0.35 &    0.00 \\
  Zn \textsc{i} &   2 &    4.18 &    0.02 & $-$0.38 &    0.30 &    0.02 &&
  Zn \textsc{i} &   2 &    3.17 &    0.04 & $-$1.39 &    0.06 &    0.03 \\
 Ba \textsc{ii} &   2 &    1.65 &    0.02 & $-$0.53 &    0.15 &    0.01 &&
 Ba \textsc{ii} &   2 &    0.78 &    0.06 & $-$1.40 &    0.05 &    0.04 \\
\cline{1-7} \cline{9-15} \\
\\
\multicolumn{7}{c}{C2306265-085103} && \\
\cline{1-7} \\
   O \textsc{i} &   2 &    7.96 &    0.02 & $-$0.72 &    0.45 &    0.02 \\
  Na \textsc{i} &   2 &    5.25 &    0.00 & $-$0.99 &    0.18 &    0.00 \\
  Mg \textsc{i} &   2 &    6.84 &    0.07 & $-$0.76 &    0.41 &    0.05 \\
  Al \textsc{i} &   4 &    5.56 &    0.08 & $-$0.89 &    0.28 &    0.04 \\
  Si \textsc{i} &   5 &    6.66 &    0.08 & $-$0.85 &    0.32 &    0.03 \\
  Ca \textsc{i} &   4 &    5.48 &    0.04 & $-$0.86 &    0.31 &    0.02 \\
 Sc \textsc{ii} &   6 &    2.18 &    0.15 & $-$0.97 &    0.20 &    0.06 \\
  Ti \textsc{i} &   4 &    4.06 &    0.03 & $-$0.89 &    0.28 &    0.01 \\
 Ti \textsc{ii} &   3 &    4.21 &    0.36 & $-$0.74 &    0.43 &    0.21 \\
   V \textsc{i} &   4 &    2.88 &    0.06 & $-$1.05 &    0.12 &    0.03 \\
  Cr \textsc{i} &   3 &    4.08 &    0.12 & $-$1.56 & $-$0.39 &    0.07 \\
 Cr \textsc{ii} &   1 &    4.39 &    0.00 & $-$1.25 & $-$0.08 &    0.00 \\
  Mn \textsc{i} &   3 &    4.35 &    0.10 & $-$1.08 &    0.09 &    0.06 \\
  Fe \textsc{i} &  62 &    6.33 &    0.12 & $-$1.17 &    0.00 &    0.01 \\
 Fe \textsc{ii} &  11 &    6.31 &    0.09 & $-$1.19 & $-$0.02 &    0.03 \\
  Co \textsc{i} &   3 &    4.00 &    0.15 & $-$0.99 &    0.18 &    0.08 \\
  Ni \textsc{i} &   7 &    5.08 &    0.08 & $-$1.14 &    0.03 &    0.03 \\
  Cu \textsc{i} &   1 &    3.55 &    0.00 & $-$0.64 &    0.53 &    0.00 \\
  Zn \textsc{i} &   2 &    3.38 &    0.14 & $-$1.18 & $-$0.01 &    0.10 \\
 Ba \textsc{ii} &   2 &    0.99 &    0.06 & $-$1.19 & $-$0.02 &    0.04
\enddata
\end{deluxetable*}

% elemental abundances

% comparison to literature for a given star

% chemical synthesis

\subsection{Distances}
% isochrone fitting

% discuss Williams et al distances

\subsection{Dynamics}

% U, V, W calculations

\section{Discussion}
\label{sec:discussion}
% wylie de boer discrepancies

\subsection{Stellar Parameter Discrepancies with \citet{wylie-de-boer;et-al_2012}}

We seek to investigate the nature of the Aquarius stream, as well as the specific globular cluster origin claim made by \citet{wylie-de-boer;et-al_2012}. The stellar parameters reported in \citet{wylie-de-boer;et-al_2012} differ slightly to those published in Table \ref{tab:stellar-parameters}. \citet{wylie-de-boer;et-al_2012} deduce their stellar parameters by minimizing the $\chi^2$ difference between the observed spectra, and the \citet{munari;et-al_2005} synthetic spectral library after it was convolved and re-sampled to match the observations. The comparison regions used for analysis were 4900-5200\,{\AA} and 5575-5725\,{\AA}. In general, temperatures between the two studies agree within the uncertainties. The only aberration is J223811-104126, where we find an effective temperature of 5140\,K, ${\sim}$500\,K cooler than the 5646\,K found by \citet{wylie-de-boer;et-al_2012}.  Similarly, \citet{williams;et-al_2011} report a hotter effective temperature of 5502\,K from low-resolution spectra. This is the largest discrepancy we see in any of our standard or program stars.

Photometric temperatures only lend marginal argumentative weight towards our spectroscopic temperatures. If the \citet{schlegel;et-al_1998} dust maps are employed, then the \citet{alonso;et-al_1999} and \citet{casagrande;et-al_2010} relations estimate the effective temperatures to be between 5104\,K and 5367\,K. Our spectroscopic temperature of 5110\,K is in excellent agreement with the \citet{alonso;et-al_1999} prediction, but not in strong disagreement with the \citet{casagrande;et-al_2010} relation either. If we employ the reddening of $E(B-V) = 0.12$ that we have determined from interstellar Na D1 line, the photometric temperature increases slightly by ${\sim}100$\,K, still within the uncertainties of the \citet{alonso;et-al_1999} scale, but less so with the \citet{casagrande;et-al_2010} scale. We find the surface gravity for J223811-104126 to be 3.00 $\pm$ 0.50, placing this star at the base of the red giant branch. Thus, if J223811-104126 is a giant star, our spectroscopic temperatures should be in slightly better agreement with the \citet{alonso;et-al_1999} scale, rather than the \citet{casagrande;et-al_2010} dwarf/sub-giant scale. However, \citet{williams;et-al_2011} and \citet{wylie-de-boer;et-al_2012} find this star to be a sub-giant, with a surface gravity $\log{g} = 4.16$ and 4.60 respectively. The \citet{williams;et-al_2011} and \citet{wylie-de-boer;et-al_2012} temperatures are 150\,K and 300\,K hotter than the \citet{casagrande;et-al_2010} prediction respectively. If our extinction value is employed, then the \citet{williams;et-al_2011} and \citet{casagrande;et-al_2010} temperatures are in good agreement, yet still 175\,K cooler than the \citet{wylie-de-boer;et-al_2012} measurement.

As a test, we set the temperature for J223811-104126 to be 5600\,K -- within the temperature regime reported by \citet{williams;et-al_2011} and \citet{wylie-de-boer;et-al_2012}. The slopes in abundance with excitation potential and reduced equivalent width were large (\#check smh: how many dex), and in doing so we could not find a physical solution for this temperature. Thus, although photometric temperatures do not help in solving this discrepancy, and the reddening measured from the interstellar Na D1 line may have made it worse, our high-resolution spectra indicates that J223811-104126 has an effective temperature of 5110\,K and is indeed a giant star.

In the \citet{wylie-de-boer;et-al_2012} study, the surface gravity was determined during the $\chi^2$ minimization process. Again, with the exception of J223811-104126, our surface gravities are largely in agreement. The only other noteworthy difference in surface gravity is for J221821-183424, where we find a lower gravity of $\log{g} = 0.80 \pm 0.30$ and \citet{wylie-de-boer;et-al_2012} find $\log{g} = 1.45 \pm 0.35$\,cm\,s$^{-2}$. Given the difference in $S/N$ and resolution between these studies, this difference is not too concerning.

\citet{wylie-de-boer;et-al_2012} calculate microturbulence from empirical relationships derived by \citet{reddy;et-al_2003} for dwarfs and \citet{fulbright_2000} for giants. These relationships are based upon the effective temperature and surface gravity. Our published microturbulent velocities agree excellently with the values presented in \citet{wylie-de-boer;et-al_2012}, again with the exception of J223811-104126, where the difference in $v_{t}$ can be directly attributed to the offsets in other observables.

Of all the stellar parameters ($T_{\rm eff}$, $\log{g}$, $v_{t}$ and $\mbox{[Fe/H]}$), the metallicities between our two studies vary the most. Here we discuss the potential causes for these discrepancies. In the \citet{wylie-de-boer;et-al_2012} study, after the stellar parameters ($T_{\rm eff}$, $\log{g}$, $v_{t}$, and an initial [Fe/H] estimate) had been determined through a $\chi^2$ analysis, the authors synthesized individual Fe \textsc{i} and \textsc{ii} lines using the \textsc{moog} code. \citet{castelli;kurucz_2003} stellar atmosphere models were employed \--- the same ones used in this study, albeit the interpolation schemes will have subtle differences. The median abundance of Fe\,\textsc{i} lines was taken as the overall stellar metallicity, scaled using the \citet{grevesse;sauval_1998} solar composition.

The study of \citet{wylie-de-boer;et-al_2012} is of slightly lower resolution ($\mathcal{R} = 25,000$ compared to $\mathcal{R} = 28,000$ presented here), but with a much lower $S/N$ ratio: $\sim{}25$ pixel$^{-1}$ compared to $>$100\,pixel$^{-1}$ achieved here. Given the slightly lower spectral resolution and modest $S/N$ in the \citet{wylie-de-boer;et-al_2012} study, there are fewer unblended absorption lines available. In fact, there are very few transitions present in their published line list: only 14 Fe\,\textsc{i} lines and 3 Fe\,\textsc{ii} lines were available. For contrast, our metallicity determinations are usually based from ${\sim}$50 Fe\,\textsc{i} and ${\sim}$13 Fe\,\textsc{ii} clean, unblended lines. Given the metallicity of these candidates and the spectral coverage (4120 to 6920 {\AA}), it is somewhat surprising that there were not more transitions available to \citet{wylie-de-boer;et-al_2012}.

The line list employed in the \citet{wylie-de-boer;et-al_2012} study contained fewer lines, and utilized different oscillator strengths than what we have used. The oscillator strengths in the \citet{wylie-de-boer;et-al_2012} study are astrophysical; they were derived from a reverse analysis from the \citet{hinkle;et-al_2003} solar atlas. Suspecting the differing line lists may be the cause of the metallicity discrepancy, we reanalysed our data using the \citet{wylie-de-boer;et-al_2012} line list and stellar parameters. Indeed, we arrive at very similar metallicities to \citet{wylie-de-boer;et-al_2012}. Without using their published stellar parameters a priori, in a `blind' analysis employing just the \citet{wylie-de-boer;et-al_2012} line list, we also arrive at similar stellar parameters. However, given the limited number of lines available, subtle changes to the stellar parameters inflicted large changes to individual mean iron abundances.  Additionally, we note that one Fe\,\textsc{i} transition at 6420.060\,\AA{} in the \citet{wylie-de-boer;et-al_2012} line list was not detected in any of our candidates at the 3-$\sigma$ level, even though the $S/N$ at this point exceeds 115 pixel$^{-1}$ in every observation. Thus, for this comparative analysis a maximum of 13 Fe\,\textsc{i} and 3 Fe\,\textsc{ii} lines were available.

Given the line list employed in the \citet{wylie-de-boer;et-al_2012} study, the overall data quality and the lack of suitable iron lines available for analysis, it appears the Aquarius stream stars conspired to present a tight metallicity distribution function of $\sigma(\mbox{Fe/H}) = 0.10$\,dex. When viewed in light of enhanced [Ni/Fe], [Na/Fe] abundance ratios, this chemistry is suggestive of a globular cluster origin. Our analysis of high-resolution spectra with high $S/N$ reveals a much broader metallicity distribution for the stream. With just five stars \--- if we assume all our candidates are members \--- we find the stream metallicity varies from $\mbox{[Fe/H]} = -0.65$ to --1.60\,dex. Although this is a small sample, we find the calculated median abundance and standard deviation to be $\mbox{[Fe/H]} = -1.26 \pm 0.32$\,dex.

\subsection{The Aquarius Stream M.D.F.}

Classical globular clusters typically exhibit very little chemical scatter. This is largely because they are primarily considered to be a single stellar populations, originating from a lone \--- largely homogenous \--- proto-stellar cloud. More recent observations have demonstrated shown that this is not always the case: many globular clusters clearly host multiple stellar populations. The separation between these populations is identifiable through chemical abundances, and often from photometry alone. In many cases, the chemistry between the two stellar populations are quite similar, still revealing minimal overall scatter within the cluster. There are some unusual exceptions to this rule, most notable of which is the globular cluster $\omega$-Centurai (NGC 5139), which is seen to host many sub-populations with very different metallicity distributions. Without the explanation of multiple sub-populations in mind, the system would be seen to exhibit an unusually wide metallicity dispersion for a globular cluster.

\begin{figure*}[t!]
	\includegraphics[width=\textwidth]{./figures/aquarius-vhel-feh.pdf}
	\caption{Helocentric velocities and metallicities for Aquarius candidates in this sample. The measurements found by \citet{williams;et-al_2011} (green squares) and \citet{wylie-de-boer;et-al_2012} (blue triangles) are shown, and stars are linked by dashed lines. \citet{wylie-de-boer;et-al_2012} used the \citet{williams;et-al_2011} velocities, which is why their measurements are coincident on the abscissa.}
	\label{fig:aquarius-vhel-feh}
\end{figure*}

Assuming all our candidates are stream members, we find a metallicity dispersion of $\sigma(\mbox{[Fe/H]}) = 0.32$\,dex. If the stream is the result of a disrupted system, this metallicity dispersion provides little indication on the progenitor. If the dispersion was smaller, confirming the observations of \citet{wylie-de-boer;et-al_2012}, a globular cluster scenario is more likely. However, given the multiple stellar populations present in many globular clusters, which can exhibit total cluster dispersions up to X dex (NGC X), Y dex (NGC Y), our dispersion of 0.32\,dex is somewhat inconvenient. Since dSph galaxies do not undergo chemical self-enrichment in the same way globular clusters do, they can exhibit significant metallicity dispersions \--- well over a dex, depending on the dwarf total luminosity. On its own, the metallicity of our stream stars does not provide any indication on the cause of the Aquarius stream, or its enrichment environment.

\subsection{The Na-O Relationship}

% establish the relationship as a true phenomena

Extensive studies looking at stars in globular clusters have revealed variations in light-element abundances, most notably an anti-correlation in sodium and oxygen content \citet[][and references therein]{norris;da_costa_1995,carretta;et-al_2009_na_o}. This chemical pattern has been repeatedly demonstrated in every well-studied globular cluster, although the strength of the anti-correlation varies in each system. Between generations of stars within a globular cluster, oxygen is converted into nitrogen, reducing the overall oxygen abundance. Even though oxygen undergoes internal mixing during its evolution along the giant branch, the detection of this chemical pattern in unevolved stars has led to the conclusion that observed abundance variations due to mixing are limited to Li, C, and N. Therefore, the light element anti-correlation required a different astrophysical explanation.

Sodium is primarily produced through through carbon burning in massive stars by the dominant $^{12}\mbox{C}(^{12}\mbox{C}, p)^{23}\mbox{Na}$ process. The final Na abundance is dependent on the neutron excess of the star, which slowly increases during carbon burning due to weak interactions \citep{arnett;truran_1974}. Massive stars ($>8 M_\odot$) eventually deliver their manufactured sodium to the interstellar-medium through Sn \textsc{ii} events. Because the eventual Sn \textsc{ii} explosion is devoid of any significant beta-decay processes, the neutron excess of the exploded material is representative of the pre-explosion abundance.

\begin{figure*}[t!]
	\includegraphics[width=\textwidth]{./figures/aquarius-o-na-cluster.pdf}
	\caption{Oxygen and sodium abundances for 14 classical globular clusters from \citet{carretta;et-al_2009a}, demonstrating the clear inverse Na-O correlation, which is ubiquitous in all well-studied globular clusters. The [O/Fe] and [Na/Fe] abundances for the five Aquarius stream candidates are also shown (last panel).}
	\label{fig:o-na-clusters}
\end{figure*}

The explosive material eventually condenses to form the next generation of stars, which will have a net increase in their neutron excess with respect to when their predecessors formed. Since the Na-production rate is correlated with the neutron excess, an overall increase in the total Na and Na-production rate between stellar generations is expected. Thus, an increase in the Na abundance between generations of massive stars is a natural consequence of nucleosynthesis. Sodium content also becomes important for production of Ni during the Sn \textsc{ii} event (see \S\ref{sec:na-ni}) because $^{23}$Na is the only stable isotope produced in significant quantities during the C- (or O-) burning stages. A small fraction of $^{22}$Na is also forged through proton capture on $^{21}$Ne in the carbon shell, an effect which is enhanced by neutrinos spallating abundant elements like $^{16}$O and $^{20}$Ne, making free protons available for capture \citep{woosley;weaver_1995}.

Oxygen depletion is likely the result of complete CNO burning within the stellar interior. The nucleosynthetic pathways that generate the Na-O anti-correlation are well known. However, the temperatures required to produce these patterns are not expected within the interiors of globular cluster stars. While the exact site for which these conditions occur remains under investigation, we can describe the abundance variation as a simple oxygen depletion (or dilution) model with time. Through comparisons with existing globular clusters, we can still draw inferences on the star-formation history of a system from measuring sodium and oxygen in a sample of its stars.

\citet{carretta;et-al_2009_na_o} defined three components of stars exhibiting an inverse Na-O correlation within a population: primordial, intermediate, and extreme components. The [Na/Fe] and [O/Fe] abundances for a given star define which component it belongs to. Primordial (first-generation) stars are those with sodium and oxygen abundances which are similar to field stars of the same metallicity. For a well-studied system, stars with the lowest cluster [Na/Fe] abundances are characterised as primordial. An intermediate component \--- the second-generation of stars \--- will have super-solar sodium abundances, and demonstrate a slight depletion in oxygen content. The oxygen exhaustion rate is ultimately dependent on the star-formation history of the cluster environment. The intermediate population is sub-divided based on how much the sodium and oxygen abundances depart from the primordial population: stars with [O/Na]$ > -0.9$ dex are classified as belonging to an extreme component of this Na-O pattern because they differ greatly from the primordial cluster abundance. The extreme O-depletion component has only been observed in a few clusters (preferentially more massive clusters with extended horizontal branches), indicating an long-lasting star-formation history in more massive clusters.

It is clear that careful consideration must be given when inferring star-formation histories from observed stellar abundances. In addition to the normal care afforded for measuring elemental abundances from high-resolution spectroscopic data, attention must be given to telluric absorption, contamination from nickel transitions, as well as non-LTE and 3D effects when calculating oxygen abundances. Furthermore, when characterising the oxygen depletion rate \--- the strength of the Na-O anti-correlation in a given cluster \--- it is vital to sample stars belonging to all three components, where possible. If only a primordial sample of stars in a globular cluster is observed, their [Na/Fe] and [O/Fe] abundances will be indistinguishable from halo stars of a similar metallicity. In such a scenario, any inferred anti-correlation could equally be explained by intrinsic chemical scatter or observational uncertainties.

% [O/Fe], [Na/Fe]
\begin{figure*}[t!]
	\includegraphics[width=\textwidth]{./figures/aquarius-o-na-halo.pdf}
	\caption{Oxygen and sodium abundances for thick disc ($+$), thin disc ($\times$) and halo ($\circ$) stars from \citet{reddy;et-al_2005}, and globular cluster stars from \citet{carretta;et-al_2009} (grey symbols). The Aquarius candidate abundances are also shown (\textbullet), illustrating how their [O/Fe], [Na/Fe] content is not dissimilar from galactic stars. Ambiguous thick disc/thin disc stars in \citet{reddy;et-al_2005} are marked with thick disc symbols ($+$), as are uncertain thick disc/halo stars with halo symbols ($\circ$).}
	\label{fig:o-na}
\end{figure*}

\citet{wylie-de-boer;et-al_2012} measured [Na/Fe] and [O/Fe] abundances for four of their six Aquarius stream members. This research has four stars in common with the \citet{wylie-de-boer;et-al_2012} study, but sodium and oxygen measurements exist for only three stars intersecting both samples, as the data quality for J223811-104126 in the \citet{wylie-de-boer;et-al_2012} study prohibited measuring oxygen. We have measurements for all of our stars, including J223811-104126, for which we ascertained an upper limit from the [O \textsc{i}] lines of [O/Fe] $< 0.5$ dex, and a corrected abundance of $\mbox{[O/Fe]} = 0.14 \pm 0.08$ from the O \textsc{i} triplet lines. In Figure \ref{fig:o-na} we have plotted theses abundances, including the corrected value for J223811-104126 rather than an upper limit.

% how do our values compare with Wylie de boer

% star 		| mine 				| hers
% C2225316	| 0.09 +/- 0.01 (2)	| 0.00 +/- 0.15
% C2306265	| 0.18 +/- 0.00 (2)| ...
% J221821	| 0.08 +/- 0.10 (1)| 0.03 +/- 0.15
% J223504	| 0.18 +/- 0.07 (3)| 0.28 +/- 0.21
% J223811	| 0.07 +/- 0.00 (1)| 0.04 +/- 0.15

The \citet{wylie-de-boer;et-al_2012} measurements indicate two stars with solar levels of Na \--- identical to field star Na abundances for their metallicity \--- and two stars with super-solar sodium content: J223504 and J232619.
We also find J223504 to be sodium-enhanced with $\mbox{[Na/Fe]} = 0.18$, slightly lower than the \citet{wylie-de-boer;et-al_2012} value of $0.28 \pm 0.21$\,dex. This discrepancy in [Na/Fe] is the largest difference between the two studies, and still within the uncertainties published by \citet{wylie-de-boer;et-al_2012}. The second star in their study with elevated sodium, J232619, is not in our sample. We find the additional star not present in the \citet{wylie-de-boer;et-al_2012} sample, C2306265, to be sodium enhanced to the same level of J223504 with $\mbox{[Na/Fe]} = 0.18$\,dex.

The sodium-enhanced stars exhibit no depletion of oxygen; their chemistry does not demonstrate a classical Na-O anti-correlation. In fact, they have a slight positive relationship: the Na-enhanced stars in the \citet{wylie-de-boer;et-al_2012} study show an increased oxygen content (within the uncertainties), contrary to an expected oxygen depletion between stellar generations within a globular cluster. With our revised oxygen abundances and two sodium-enhanced stars, we find a very slight positive relationship of X.XX +/- X.XX, and no evidence of a Na-O anti-correlation in the Aquarius stream.

If the Aquarius stream is the result of a disrupted globular cluster, a large part of the picture would still be missing. Most of the Aquarius stream stars studied to date (either in this sample or the \citet{wylie-de-boer;et-al_2012} study), would be classified as belonging to a `primordial' component, with chemistry indistinguishable from the halo. Identifying more Aquarius stream members belonging to the intermediate component with strong oxygen depletion, or perhaps members of an extreme component, would be strong evidence for a Na-O anti-correlation and perhaps a globular cluster origin. Three stream stars identified to date (including two from this sample) could plausibly be classified as members of an intermediate population, with only a slight enhancement in sodium and no oxygen-depletion. Thus, if the strength of any Na-O relationship is to be used to vet potential disrupted hosts for the Aquarius stream, many more stream candidates will need to be identified, and observed with high-resolution and high $S/N$.

\subsection{The Al-Mg Relationship}

%The Na-O anti-correlation is likely the result of complete CNO and Ne-Na cycles of proton reactions in globular cluster stars. There is also evidence in some clusters that the Mg-Al nucleosynthesis cycle is also active.

An inverse relationship between sodium and oxygen is not the only chemical pattern commonly observed in globular clusters. Although not ubiquitous to every system, many clusters exhibit an anti-correlation between aluminium and magnesium abundances. This is perhaps unsurprising, given the nucleosynthetic pathways for aluminium and magnesium. In addition to the CNO cycle operating during hydrogen burning, the Mg-Al chain can also operate under extreme temperatures \citep[$T_6 \sim 3$\,K;][]{arnould;et-al_1999}. Aluminium is produced by proton capture of magnesium, beginning with $^{24}$Mg to $^{25}$Al. The lifetime of $\beta$-decay to proton capture allows for the production of unstable $^{27}$Si through proton capture. Seconds later, the isotope decays to $^{27}$Al, completing the $^{27}$Si path of the Mg-Al cycle. The alternative process from $^{25}$Al involves proton capture to $^{26}$Mg, but this process requires much higher temperatures ($T_6 \sim 7$\,K). In fact, although the process operates from temperatures above 30 million K, large Mg-Al abundance variations are not appreciable unless the temperatures reach $T_6 \sim 6.5$\,K.

The Mg-Al cycle can explain the observed anti-correlation observed in some globular clusters, and most theoretical models predict such a pattern (\#get citations from Amanda). However, the locations where these extreme temperatures can be found are under some debate. Theoretical models indicate that the interiors of globular cluster stars do not reach sufficient temperatures for the Mg-Al cycle to produce the observed variations. Thus, although the processes for creating these chemical patterns are understood, the exact site and requisite conditions are lacking a full description.

No inverse correlation has been observed in the Aquarius stream stars to date. Given that we do not find a Na-O anti-correlation, it is not surprising that a Mg-Al pattern has also not been observed. \citet{wylie-de-boer;et-al_2012} published magnesium and aluminium abundances for five stars in their Aquarius stream sample. No inverse correlation is present, and their abundances are not dissimilar from galactic stars. The abundances tabulated in Table \ref{tab:abundances} vary slightly with those presented in the \citet{wylie-de-boer;et-al_2012} sample. Two candidates are in good agreement: J221821 and J223811, albeit the data quality in the \citet{wylie-de-boer;et-al_2012} study did not permit measuring aluminium for J223811. The other two stars common between studies are C2225316 and J223504, which we found to be the most abundant in both magnesium and aluminium.

While \citet{wylie-de-boer;et-al_2012} find almost no scatter ($\pm0.02$\,dex) in [Mg/Fe] for stars common to both studies, we find C2225316 and J223504 to be almost $+0.20$\,dex higher than the rest of the sample. The reason for this discrepancy is not obvious, given the stellar parameters between the two studies agree well, with the previously discussed exception of metallicity. In particular, the stellar parameters agree excellently for J223504, but we find $\mbox{[Fe/H]} = -0.68$ instead of the $-0.98$ as found by \citet{wylie-de-boer;et-al_2012}. Of the Mg \textsc{i} profiles measured, only two transitions are common to both line lists: 6318$\lambda$ and 6319$\lambda$. The oscillator strengths differ between studies; in these two lines the $\log{gf}$ differs by $-0.24$ and $-0.27$\,dex respectively (our oscillator strengths are lower). This discrepancy only complicates the matter further, since if all other things being equal, then our abundances ought to be lower than the \citet{wylie-de-boer;et-al_2012} study.

% how do our values compare with Wylie de boer
%							[Mg/Fe]						[Al/Fe]					| [Mg?]	| [Al?]
% star 		| mine 				| hers			| mine			| hers			| 
% C2225316	| 0.50 +/- 0.06		| 0.32 +/- 0.05	| 0.68 +/- 0.04	| 0.38 +/- 0.12	| No, + | No, +
% C2306265	| 0.41 +/- 0.05		| ...			| 0.28 +/- 0.04	| ...			| N/A	| N/A 
% J221821	| 0.34 +/- 0.05		| 0.30 +/- 0.01	| 0.21 +/- 0.10	| 0.29 +/- 0.12	| Yes +	| Yes -
% J223504	| 0.49 +/- 0.09		| 0.33 +/- 0.05	| 0.31 +/- 0.05	| 0.21 +/- 0.03	| No +	| No +
% J223811	| 0.32 +/- 0.01		| 0.30 +/- 0.06	| 0.10 +/- 0.10	| ...			| Yes +	| N/A

% Mg-Al
\begin{figure}[h!]
	\includegraphics[width=\columnwidth]{./figures/aquarius-mg-al.pdf}
	\caption{Magnesium and aluminium abundances for Aquarius candidates, and Milky Way halo/disc stars from \citet{reddy;et-al_2003} ($+$) and \citet{fulbright_2000} ($\circ$). The chemically peculiar star, C2225316-14437, is marked.}
	\label{fig:mg-al}
\end{figure}

The [Mg/Fe] and [Al/Fe] abundances found for the five Aquarius stars in this sample are illustrated in Figure \ref{fig:mg-al}. Interestingly, our stars exhibit a strong positive relationship, with a best-fitting slope of $\mbox{[Al/Fe]} = 2.15 x \mbox{[Mg/Fe]} - 0.57$ ($R = Y.YY$). If the exclude C2225316, our chemically peculiar star with enhanced [Al/Fe] relative to the rest of the sample, the slope decreases to $\mbox{[Al/Fe]} = 1.07 x \mbox{[Mg/Fe]} - 0.19$ and the correlation coefficient $R$ increases to Z.ZZ. Even when a Mg-Al anti-correlation is not detected in globular clusters, there is generally some scatter in [Al/Fe] at constant [Mg/Fe] \citet[e.g. see][]{carretta;et-al_2009a}. Few classical globular clusters exhibit a positive correlation.

% Mg-Al
\begin{figure*}[t!]
	\includegraphics[width=\textwidth]{./figures/aquarius-mg-al-cluster.pdf}
	\caption{[Mg/Fe] and [Al/Fe] abundances for stars in 16 globular clusters from \citet{carretta;et-al_2009} and for the five Aquarius stream candidates. The inverse Mg-Al correlation is illustrated, most obviously in NGC 2808 and NGC 1904. The strong positive correlation observed in Aquarius is in stark contrast with other classical globular clusters.}
	\label{fig:mg-al}
\end{figure*}

A positive relationship between magnesium and aluminium can result from SN \textsc{ii} contributions to the local interstellar medium. High-resolution observations of disc stars illustrate this correlation \citep{bensby;venn}, demonstrating efficient SN \textsc{ii} mixing. Some intermediate-mass ($4M_\odot$) AGB models can also predict a positive correlation between aluminium and magnesium \citet{karakas;et-al}. Under extreme temperatures ($T_6 \sim 300$\,K), substantial $^{25}$Mg and $^{26}$Mg are produced by $\alpha$-capture onto $^{22}$Ne by the $^{22}$Ne($\alpha$,n)$^{25}$Mg and $^{22}$Ne($\alpha$,$\gamma$)$^{26}$Mg reactions respectively. Following deep dredge up, significant $^{25}$Mg and $^{26}$Mg is mixed into the photosphere, more than the quantity of $^{26}$Al produced through the Mg-Al cycle. This dredge up process replenishes the Mg which is expensed during the Mg-Al pathway. Therefore, a positive relationship between magnesium and aluminium can occur if there has been significant contributions from intermediate mass AGB stars undergoing dredge up.

An unrealistically high fraction of intermediate-mass AGB stars would be required to produce the same chemical signature of a single Sn \textsc{ii} event. Given the efficiency of chemical pollution following supernovae, the observed correlation in the thick disc is likely the result of Sn \textsc{ii} events. However, since the chemical enrichment environment of the Aquarius stream is unknown, we cannot exclude either possibility. Distinguishing between these processes observationally requires careful measurements of magnesium isotope abundances $^{24}$Mg (indicating supernovae mixing), $^{25}$Mg and $^{26}$Mg (suggesting significant AGB contribution), which is not possible given our $S/N$ or spectral resolution.

\subsection{The Na-Ni Relationship}

% need to say about [alpha/fe] and [na/fe] with distance
Detailed chemical studies of nearby disk stars have noted a correlation with [Na/Fe] and [Ni/Fe] abundances. This relationship was first hinted in \citet{nissen;schuster_1997}, where the authors found eight stars that were underabundant in [$\alpha$/Fe], [Na/Fe] and [Ni/Fe]. Interestingly, the authors noted that stars at larger galactocentric radii were most deficient in these elements. \citet{fulbright_2000} saw a similar signature: stars with low [Na/Fe] were only found at large ($R_\odot > 20$\,kpc) distances. \citet{nissen;schuster_1997} proposed that since the outer halo is thought to have been built up by the accretion of dwarf satellite galaxies, then the Na-Ni pattern may be a chemical indicator of merger history within the galaxy.

With additional data from \citet{nissen;schuster_2011}, the Na-Ni relationship was found to be slightly steeper than originally put forward. The pattern exists only for stars with $-1.5 < \mbox{[Fe/H]} < -0.5$ dex, and is not seen in metal-poor dSph stars, providing a potentially useful indicator for investigating chemical evolution.

%However, it is crucial to note that although there are only a few dSph stars in the $-1.5 < \mbox{[Fe/H]} < -0.5$ metallicity regime with [Na/Fe], [Ni/Fe] measurements, they agree reasonably well with the galactic trend.

The correlation between sodium and nickel content is the nucleosynthetic result of neutron capture in massive stars. As previously discussed, the total Na abundance is controlled by the neutron excess, which limits the production of $^{58}$Ni during Sn \textsc{ii} events. When the inevitable supernova begins, the core photodissociates into neutrons and protons, allowing the temporary creation of $^{56}$Ni before it decays to $^{56}$Fe. A limited amount of $^{54}$Fe is also formed, which is the main source of production for the stable $^{58}$Ni isotope through $\alpha$-capture. When the core dissociates, the quantity of $^{54}$Fe (and hence $^{58}$Ni) produced is dependent on the abundance of neutron-rich elements during the explosion. As $^{23}$Na is a relatively plentiful neutron source with respect to other potential sources (like $^{13}$C), the post-supernova $^{58}$Ni abundance is driven by the pre-explosion $^{23}$Na content. Recall that the total Na production primarily depends on the neutron excess. Thus, through populations of massive stars undergoing C-burning, a positive correlation between sodium and nickel can be expected.

%As \citet{Venn;et-al_2004} point out, this pattern can be altered from contributions by type Ia supernova.

% [Na/Fe], [Ni/Fe]
\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./figures/aquarius-na-ni.pdf}
	\caption{A plot showing [Na/Fe] and [Ni/Fe] for all Aquarius stream stars. TODO add dwarf/GC/disk stars on here.}
	\label{fig:na-ni}
\end{figure}

Stars originating in dwarf spheroidal galaxies and globular clusters have very different chemical enrichment environments. Consequently, they exhibit chemistry that reflects their nucleosynthetic antiquity. Dwarf spheroidal stars do not demonstrate enhanced sodium or nickel content with respect to iron, as there has been a relatively negligible linage of massive stars undergoing supernova. In contrast, globular cluster stars do have elevated [Na/Fe], [Ni/Fe] signatures. This sharp contrast between dwarf spheroidal and globular cluster star chemistry is illustrated in Figure \ref{fig:na-ni}. Given the extended star formation within the Milky Way disk, globular cluster stars and disk stars are indiscernible in the Na/Ni plane: they both show an extended contribution of massive stars.

The most that can be inferred from the Na, Ni abundances of Aquarius stream stars is that their enrichment environment is less like a dSph galaxy, and more representative of a globular cluster, or the disc of the Milky Way.

\subsection{The Chemical Distinctiveness of C2225316}

% for a high mass AGB
%Sr/Ba ~ 0.4

C2225316 exhibits peculiar chemical abundances compared to the rest of the Aquarius candidates. It is over-abundant in light refractory elements, and exhibits a strong barium abundance of $\mbox{[Ba/Fe]} = 0.70$\,dex. This abundance is well in excess of the halo ($\mbox{[Ba/Fe]} \sim 0.0$) \--- and our other Aquarius stream stars \--- which vary between $-$0.02 to 0.15\,dex.

High $s$-process enhancements have been noted for a few disrupted globular clusters that exhibit a wide metallicity range. Most noteably: $\omega$Centurai \citep{majewski;et-al_2012}, M22 \citep{marino;et-al_2011} and (\# any others?). Like the Aquarius co-moving group, both clusters are relatively close to the sun: $R_\odot = 5.2$\,kpc and 3.2\,kpc, respectively. M22 has a mean metallicity of $\mbox{[Fe/H]} \sim -1.7$, varying between $-2.0 < \mbox{[Fe/H]} < -1.6$\,dex. Since C2225316 has a metallicity of $\mbox{[Fe/H]} = -1.26$\,dex, it is unlikely to be associated with M22. However, the high barium content and overall metallicity of C2225316 ($\mbox{[Fe/H]} = -1.26$) suggests there may be an association with $\omega$-Centurai.

$\omega$-Centurai has a retrograde orbit with low inclination. Many groups simulating this orbit have predicted retrograde tidal debris to occur near the Solar circle \citep{dinescu_2002,tsuchiya_2003,tsuchiya_2004,mizutani_2003,bekki;freeman_2003}. Subsequent searches for $\omega$-Cen debris in the solar neighbourhood have led to tantalizing signatures of debris. From over 4,000 stars selected from the $\omega$-Cen color-magnitude diagram by \citet{da_costa;coleman_2008}, only six candidate members were recovered. More recently, \citet{majewski;et-al_2012} presented data from the Grid Giant Star Survey (GGSS), an all-sky search looking for metal-poor giant stars. From their data, they identified 12 stream candidates. Although this scarcity is consistent with tidal stripping occurring long ago, it is perhaps worrying that searches for stripped stars near $\omega$-Cen itself have been even less fruitful.

In addition to presenting new stream candidates, \citet{majewski;et-al_2012} performed 4,050 different simulations in order to predict likely locations for tidal debris. The results of their simulation are replicated in Figure \ref{fig:majewski-simulation}, where the location of C2225316 is also shown. Interestingly, the velocity and position of C2225316 line up almost precisely where \citet{majewski;et-al_2012} predict a high probability of stream members.

Although this is interesting from a qualitative perspective, the most striking evidence for an association with $\omega$-Centurai is in chemistry. Figure \ref{fig:omega-cen-barium} demonstrates the [Ba/Fe] abundance for $\omega$-Cen stars compared to that of the galactic halo. A steep Ba enhancement is present, which is unique to the globular clusters $\omega$-Cen and M22. We note that although high $s$-process enrichment is present in other systems \citep[IC 2391, the Argus association;][]{de_silva;et-al_2013}, these systems are have metallicities near solar values and are unlikely to be associated with C2225316.

% additional chemistry?
% [Sr/Ba] ~0.4

The chemical peculiarity of C2225316 suggests that it is distinct from the Aquarius co-moving group. The high $s$-process enrichment is similar to that observed in $\omega$-Centurai stars, and the overall metallicity ($\mbox{[Fe/H]} = -1.26$) agrees excellently with the mean metallicity of a known sub-population within $\omega$-Centurai. Moreover, the position and velocity for C2225316 precisely coincide with the most likely predictions for tidal debris from $\omega$-Centurai. The chemical and phase-space information indicates that C2225316 is likely associated with the disrupted globular cluster $\omega$-Centurai.

\subsection{Possible Progenitors}

Since the Aquarius stream is kinematically coherent, the default assumption is that the system has been accreted. The stream candidates presented here are slightly enhanced in sodium and nickel, suggesting that if a progenitor ever existed then its environment was more like a globular cluster than a dSph galaxy. Based on a lack of $\alpha$-enhancement and increased [Na/Fe], [Ni/Fe] abundances, we exclude the possibility that the Aquarius stream has resulted from a disrupted dwarf satellite. If the progenitor is a disrupted globular cluster, and all our candidates are members, then it has a wider metallicity distribution than most globular clusters.

However, our chemical analysis does not reveal any inverse Na-O correlation \--- which would be strong evidence for a globular cluster environment. Furthermore, we observe a positive Mg-Al relationship, the same seen by \citet{wylie-de-boer;et-al_2012}. High resolution, high $S/N$ spectra for the entire stream may be required before such chemical signatures may become apparent. In this section we discuss other possible causes for the Aquarius co-moving group.

\subsubsection{$\omega$-Centurai}

In their original discovery paper, \citet{williams;et-al_2011} attempted to exclude possible known progenitors. On the basis of metallicity, distance, proper motions\footnote{We find the proper motions in Table 1 of \citet{williams;et-al_2011} to be mis-associated to incorrect stars. However, this  error was typographical; it did not affect the transverse velocity calculations (Williams, M.E.K, private communication).}, transverse velocities and orbital energies, the authors were able to exclude all known Milky Way satellites, with the notable exception of $\omega$-Centurai. Although the Aquarius stream metallicity distribution we find is not dissimilar from a known sub-population in $\omega$-centurai, the individual chemical abundances are quite distinct. $\omega$-Centurai members exhibit a strong $s$-process enhancement with overall metallicity. This signature is not observed in our data, with the exception of C2225316, which seems to be chemically distinct from the other Aquarius group candidates.

Due to our limited sample size, it is impossible to know whether C2225316 is an outlier, or our four other candidates are halo interlopers. Either position is debatable, and both lead to interesting interpretations. If most of our observed candidates are halo members, and C2225316 represents the Aquarius stream chemistry, then an association between Aquarius and $\omega$-Centurai is possible.

%Alternatively, the stream may indicate first evidence of a minor merger within the Milky Way disc. However, with certainty we can state that most observed Aquarius stream candidates do not exhibit enhanced $s$-process elements, and are therefore unlikely to be associated with $\omega$-Centurai \--- or any other globular cluster.

\subsubsection{Disrupted Disk Stars \--- Signature of a Minor Merger}

\# Velocity distribution wide

\# Spatial distribution wide

\# Chemistry rules out a dSph origin

\# No signs of classical GC chemical patterns

\# The stream chemistry is indistinguishable from disk stars in every aspect

\# Revised distances to each star, transverse motions + orbital energies calculated

\# Monte-carlo simulations demonstrate strongly retrograde stars wrt thick disk + halo

\# Orbital energy pattern aligns exactly with Zolotov+ simulations, but would require a smaller mass of the disrupter (minor merger)

\# Need to discuss disk is unlikely to have had a merger in past 8 Gyr

% could they be displaced thick disk stars?

% [Na/Fe], [Cr/Fe], [Ni/Fe]
\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./figures/aquarius-na-cr-ni-fe.pdf}
	\caption{A plot showing [Na/Fe], [Cr/Fe], and [Ni/Fe] for all Aquarius stream stars.}
	\label{fig:na-cr-ni-fe}
\end{figure}

% Toombre #TODO - make this a 4panel in UVW
\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./figures/plot-toombre.pdf}
	\caption{Toombre plot.}
	\label{fig:toombre}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}
\# Conclusions

%\facilities{Magellan:Clay}

\bibliographystyle{apj}
\bibliography{bibliography}

\end{document}