### 2013-02-27 thesis/appendices/appendices.tex

\chapter{An image of a kitten}

A kitten is a juvenile domesticated cat. A feline litter usually consists of two to five kittens. To survive, kittens need the care of their mother for the first several weeks of their life. Kittens are highly social animals and spend most of their waking hours playing and interacting with available companions. The word \emph{``kitten''} derives from Middle English \emph{kitoun} (ketoun, kyton etc.), which itself came from Old French \emph{chitoun}. The young of big cats are called cubs rather than kittens. Either term may be used for the young of smaller wild felids such as ocelots, caracals, and lynx, but \emph{kitten} is usually more common for these species.

\begin{figure}[b!] 
   \centering
   \includegraphics[width=0.9\textwidth]{kitten.jpg} 
   \caption{The aforementioned kitten}
   \label{fig:kitten}
\end{figure}

\chapter{Lots of awful mathematics}
\input{epicyclic}