### 2013-02-27 thesis/disclaimer.tex

\section*{Disclaimer}

I hereby declare that the work in this thesis is that of the candidate
alone, except where indicated below or in the text of the thesis. The work was undertaken between February 2010 and July 2013 at the Australian National University, Canberra, except for the period between July 2012 and December 2012 where the work was undertaken at the Massachusetts Institute of Technology, Boston. It has not been submitted in whole or in part for any other degree at this or any other university.

\vspace{1cm}
\begin{flushright}
\includegraphics[width=0.4\linewidth]{signature.png}\\[2mm]
\fullname

\fullthesisdate
\end{flushright}